## Agency Arrests Over Time

```{r 'analysis_agencyCounts_byAgency', echo = FALSE}
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_agencyCounts.pdf",
                                       "analysis_agencyCounts_agencyCourtDispositions.pdf"))
if (any(!file.exists(imgFiles))) {
  remapFactor <- function(x, groups, name, droplevels = TRUE) {
    x[x %in% groups[-1L]] <- groups[1L]
    levels(x)[levels(x) == groups[1L]] <- name
    if (droplevels) x <- base::droplevels(x)
    x
  }
  
  loadData(mainTable, "..")
  
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                c("acquited", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                  "diversion_dismissed", "not_guilty_insane"),
                "crt",
                FALSE)
  
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                  "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified"),
                "lea",
                FALSE)
  
  obts.sub$disposition <- 
    remapFactor(obts.sub$disposition,
                c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                  "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                  "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                  "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable"),
                "da",
                FALSE)
  obts.sub$disposition <- droplevels(obts.sub$disposition)
  rm(remapFactor)
  
  years <- sort(unique(obts.sub$court_event_year))
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  
  totalsByJurisdiction <- apply(totals, 1L, sum)
  
  cutoffs <- seq(log(max(totalsByJurisdiction) + 1), 0, length.out = 10L)[-1L]
  
  indexSets <- list()
  indexSets[[1L]] <- unname(which(log(totalsByJurisdiction + 1) > cutoffs[1L]))
  for (i in seq.int(2L, length(cutoffs))) indexSets[[i]] <- unname(which(cutoffs[i] < log(totalsByJurisdiction + 1) & log(totalsByJurisdiction + 1) <= cutoffs[i - 1L]))
  
  rm(totalsByJurisdiction, cutoffs)
  
  pdf(imgFiles[1L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, max(totals[indices,])),
         xlab = "year", ylab = "count", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
    plotRange <- par()$usr
    xVal <- 0.975 * (plotRange[2L] - plotRange[1L]) + plotRange[1L]
    yVal <- 0.975 * (plotRange[4L] - plotRange[3L]) + plotRange[3L]
    text(xVal, yVal, paste0("n = ", length(indices)), adj = c(1, 0.75))
  }
  dev.off()
  
  pdf(imgFiles[2L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, 1),
         xlab = "year", ylab = "% crt", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, dispoTable[indices[j],,"crt"] / totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
  }
  dev.off()

  rm(i, j, indices, xVal, yVal, plotRange)
  
  rm(years, indexSets, dispoTable, totals)
}
rm(imgFiles)
```

This section shows the total number of dispositions by arresting agency over time, as well as the percentage of records that have a court disposition. Since there are a few agencies that account for most of the arrests, the break points in the following plots are made uniform over the log of the number of arrests, plus 1.

#### Agency Totals

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_agencyCounts.pdf"))`

`r rmdPageBreak()`

#### Agency Court Disposition Percentages

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_agencyCourtDispositions.pdf"))`

`r rmdPageBreak()`

## County Arrests Over Time

```{r 'analysis_agencyCounts_byCounty', echo = FALSE}
txtFile <- file.path("..", txtPath, c("analysis_agencyCounts_lowCounties.txt"))
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_countyCounts.pdf",
                                       "analysis_agencyCounts_countyCourtDispositions.pdf"))
if (!file.exists(txtFile) || any(!file.exists(imgFiles))) {
  remapFactor <- function(x, groups, name, droplevels = TRUE) {
    x[x %in% groups[-1L]] <- groups[1L]
    levels(x)[levels(x) == groups[1L]] <- name
    if (droplevels) x <- base::droplevels(x)
    x
  }
  
  loadData(mainTable, "..")
  
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                c("acquited", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                  "diversion_dismissed", "not_guilty_insane"),
                "crt",
                FALSE)
  
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                  "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified"),
                "lea",
                FALSE)
  
  obts.sub$disposition <- 
    remapFactor(obts.sub$disposition,
                c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                  "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                  "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                  "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable"),
                "da",
                FALSE)
  obts.sub$disposition <- droplevels(obts.sub$disposition)
  rm(remapFactor)
  
  obts.sub$county <- as.factor(substr(as.character(obts.sub$arresting_agency), 1L, 2L))
  obts.sub <- obts.sub[,c("county", "court_event_year", "disposition")]
  
  years <- sort(unique(obts.sub$court_event_year))
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  
  totalsByCounty <- apply(totals, 1L, sum)
  
  
  loadData("judicialDistricts", "..")
  countyNameMap <- unique(judicialDistricts[,c("county_code", "county")])
  percentages <- apply(dispoTable, c(1L, 2L), function(x) x["crt"]) / totals
  
  lowCounties <- which(apply(percentages, 1L, function(row) any(row <= 0.4, na.rm = TRUE)))
  lowCountyInfo <- data.frame(code = names(lowCounties), name = countyNameMap$county[match(names(lowCounties), countyNameMap$county_code)])
  lowCountyInfo$years <- sapply(lowCounties, function(index) {
    paste0(as.character(years[which(percentages[index,] <= 0.4)]), collapse = ", ")
  })
  lowCountyInfo$ave_yearly_count <- round(totalsByCounty[lowCounties] / length(years), 0)
  lowCountyInfo <- lowCountyInfo[order(lowCountyInfo$ave_yearly_count, decreasing = TRUE),]
  lowCountyInfo <- rmdFormat(lowCountyInfo, maxRows = NA)
  rm(countyNameMap, lowCounties)
  
  cutoffs <- c(1000000L, 250000L, 100000L, 25000L, 10000L, 1L)
  
  indexSets <- list()
  indexSets[[1L]] <- unname(which(totalsByCounty > cutoffs[1L]))
  for (i in seq.int(2L, length(cutoffs))) indexSets[[i]] <- unname(which(cutoffs[i] < totalsByCounty & totalsByCounty <= cutoffs[i - 1L]))
  
  rm(totalsByCounty, cutoffs)
  
  pdf(imgFiles[1L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, max(totals[indices,])),
         xlab = "year", ylab = "count", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
    plotRange <- par()$usr
    xVal <- 0.975 * (plotRange[2L] - plotRange[1L]) + plotRange[1L]
    yVal <- 0.975 * (plotRange[4L] - plotRange[3L]) + plotRange[3L]
    text(xVal, yVal, paste0("n = ", length(indices)), adj = c(1, 0.75))
  }
  dev.off()
  
  pdf(imgFiles[2L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, 1),
         xlab = "year", ylab = "% crt", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, percentages[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
  }
  dev.off()

  rm(i, j, indices, xVal, yVal, plotRange)
  
  rm(years, indexSets, dispoTable, totals, percentages)
  
  writeLines(lowCountyInfo, txtFile)
} else {
  lowCountyInfo <- readLines(txtFile)
}
rm(txtFile, imgFiles)
```

Here, we replicate the above plots but first aggregate arresting agencies into their counties.

#### Count Totals

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_countyCounts.pdf"))`

`r rmdPageBreak()`

#### County Court Disposition Percentages

Sharp dips in the arrests associated with a court disposition are likely due to reporting problems at the DA or court level. 

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_countyCourtDispositions.pdf"))`

Those counties with court disposition percentages that dip below 40% are:

`r paste0(lowCountyInfo, collapse = "\n")`

```{r echo = FALSE}
rm(lowCountyInfo)
```
