## Agency Arrests Over Time

```{r 'analysis_agencyCounts_agencyTotals', echo = FALSE}
txtFile <- file.path("..", txtPath, "analysis_agencyCounts_agencyTotals.txt")
imgFile <- file.path("..", imgPath, "analysis_agencyCounts_agencyTotals.pdf")
if (!file.exists(txtFile) || !file.exists(imgFile)) {
  loadData(mainTable, "..")
  
  numAgencies <- nlevels(obts$arresting_agency)
  numYears <- length(unique(obts.sub$court_event_year))
  
  writeLines(as.character(c(numAgencies, numYears)),
             txtFile)
  
  agencyTable <- table(obts$arresting_agency)
  
  pdf(imgFile, 6, 3)
  par(mfrow = c(1L, 2L))
  hist(agencyTable,
       breaks = 25L, freq = TRUE, xlab = "Num Arrests",
       main = "Agency Histogram")
  hist(evalx(agencyTable, x[x < quantile(x, 0.95)]),
       breaks = 25L, freq = TRUE, xlab = "Num Arrests",
       main = "Agency Histogram Lower 95%")
  dev.off()
  
  rm(agencyTable)
} else {
  lines <- readLines(txtFile)
  numAgencies <- as.integer(lines[1L])
  numYears <- as.integer(lines[2L])
  rm(lines)
}
rm(txtFile, imgFile)
```

This section attempts to show the distribution of the number of arrests as a function of agency as time. As there are `r numAgencies` agencies and `r numYears` years this involves a lot of data, so we settle first for giving a rough presentation before picking a few anomalous agencies to examine in more detail. However, first we show the distribution of the number of arrests for each agency by aggregating over time.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_agencyTotals.pdf"))`

```{r echo = FALSE}
rm(numAgencies, numYears)
```

```{r 'analysis_agencyCounts_byAgency', echo = FALSE}
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_agencyCounts.pdf",
                                       "analysis_agencyCounts_agencyCourtDispositions.pdf"))
if (any(!file.exists(imgFiles))) {
  loadData(mainTable, "..")
  
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  courtDispositions <- c("acquitted", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                         "diversion_dismissed", "not_guilty_insane")
  releasedDispositions <- c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                            "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified")
  declinedDispositions <- c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                            "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                            "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                            "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable")
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                list(courtDispositions, releasedDispositions, declinedDispositions),
                c("crt", "lea", "da"))
  rm(courtDispositions, releasedDispositions, declinedDispositions)
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  
  totalsByAgency <- apply(totals, 1L, sum)
  
  cutoffs <- seq(log(max(totalsByAgency) + 1), 0, length.out = 10L)[-1L]
  
  indexSets <- list()
  indexSets[[1L]] <- unname(which(log(totalsByAgency + 1) > cutoffs[1L]))
  for (i in seq.int(2L, length(cutoffs))) indexSets[[i]] <- unname(which(cutoffs[i] < log(totalsByAgency + 1) & log(totalsByAgency + 1) <= cutoffs[i - 1L]))
  
  rm(totalsByAgency, cutoffs)
  
  years <- sort(unique(obts.sub$court_event_year))
  
  pdf(imgFiles[1L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, max(totals[indices,])),
         xlab = "year", ylab = "count", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
    plotRegion <- par()$usr
    xVal <- 0.975 * (plotRegion[2L] - plotRegion[1L]) + plotRegion[1L]
    yVal <- 0.975 * (plotRegion[4L] - plotRegion[3L]) + plotRegion[3L]
    text(xVal, yVal, paste0("n = ", length(indices)), adj = c(1, 0.75))
  }
  dev.off()
  
  pdf(imgFiles[2L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, 1),
         xlab = "year", ylab = "% crt", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, dispoTable[indices[j],,"crt"] / totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
  }
  dev.off()

  rm(i, j, indices, xVal, yVal, plotRegion)
  
  rm(years, indexSets, dispoTable, totals)
}
rm(imgFiles)
```

### Agency Totals

In what follows, we plot individual time trends for all of the agencies. Because of the wide spread in ranges for the number of arrests per agency, we divide the trends up based on the logarithm of the total number of arrests per agency (across all years), plus 1. Binning the jurisdictions uniformly across this range produces a roughly equivalent number of in each plot, with the expection of those making the most arrests. `n` refers to the number of jursidictions being displayed. What is important to note is that some agencies show very sharp increases and decreases in the number of arrests.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_agencyCounts.pdf"))`

`r rmdPageBreak()`

#### Agency Court Disposition Percentages

Here we produce plots similar to those above, but instead look at the percentage of arrests that have a court disposition, i.e. were not released by law enforcement or declined to prosecute by the DA. The agencies in each grid square correspond to those in the previous plot, and again it is the sharp increases and decreases that are of interest. Also concerning are values near 0 or 1 when the number of arrests are large, as they may potentially indicate under-reporting.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_agencyCourtDispositions.pdf"))`

`r rmdPageBreak()`

### Agencies with Large Jumps

```{r 'analysis_agencyCounts_anomalousAgencies', echo = FALSE}
txtFile <- file.path("..", txtPath, "analysis_agencyCounts_anomalousAgencies_largeAgencies.txt")
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_anomalousAgencies_numArrestIncreases.pdf",
                                       "analysis_agencyCounts_anomalousAgencies_numArrestDecreases.pdf",
                                       "analysis_agencyCounts_anomalousAgencies_dispositionIncreases.pdf",
                                       "analysis_agencyCounts_anomalousAgencies_dispositionDecreases.pdf",
                                       "analysis_agencyCounts_anomalousAgencies_dispositionMaxes.pdf",
                                       "analysis_agencyCounts_anomalousAgencies_dispositionMins.pdf"))
if (!file.exists(txtFile) || !all(file.exists(imgFiles))) {
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  courtDispositions <- c("acquitted", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                         "diversion_dismissed", "not_guilty_insane")
  releasedDispositions <- c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                            "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified")
  declinedDispositions <- c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                            "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                            "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                            "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable")
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                list(courtDispositions, releasedDispositions, declinedDispositions),
                c("crt", "lea", "da"))
  rm(courtDispositions, releasedDispositions, declinedDispositions)
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  totalsByAgency <- apply(totals, 1L, sum)
  
  cdf <- cumsum(sort(totalsByAgency)) / sum(totalsByAgency)
  agencyCutoff <- totalsByAgency[order(totalsByAgency)[which.min(cdf < 0.02) - 1L]]
  agenciesToKeep <- totalsByAgency > agencyCutoff
  rm(cdf, agencyCutoff)
  
  
  arrestingAgencies <- loadData("arrestingAgencies", "..")
  years <- sort(unique(obts.sub$court_event_year))
  
  numAgenciesKept <- sum(agenciesToKeep)
  writeLines(as.character(numAgenciesKept), txtFile)
} else {
  numAgenciesKept <- as.character(readLines(txtFile))
}
```

Next we single out agencies with the largest increases and decreases year over year, excluding those years with 0 arrests. We focus only on those agencies responsible for 98% of arrests, or the top `r numAgenciesKept` agencies.

```{r, echo = FALSE}
rm(numAgenciesKept)
```

```{r 'analysis_agencyCounts_anomalousAgencies_numArrests', echo = FALSE}
if (!file.exists(txtFile) || !all(file.exists(imgFiles))) {
  ratios <- apply(totals[agenciesToKeep,], 1L, function(row) evalx(tail(row, -1L) / head(row, -1L), x[is.finite(x) & x > 0]))
  increases <- sapply(ratios, function(x) if (length(x) > 0L) max(x) else 0)
  decreases <- sapply(ratios, function(x) if (length(x) > 0L) min(x) else 1)
  rm(ratios)
  
  ratioOrder <- order(increases, decreasing = TRUE)
  
  pdf(imgFiles[1L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, totals[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         main = agencyName, cex.main = 0.7,
         ylab = "num arrests", xlab = "year")
  }
  dev.off()
  
  ratioOrder <- order(decreases)
  
  pdf(imgFiles[2L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, totals[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         main = agencyName, cex.main = 0.7,
         ylab = "num arrests", xlab = "year")
  }
  dev.off()
  
  rm(i, agency, agencyName, ratioOrder, increases, decreases, totals)
}
```

#### Largest Increases - Agency Number of Arrests

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_numArrestIncreases.pdf"))`

`r rmdPageBreak()`

#### Largest Decreases - Agency Number of Arrests

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_numArrestDecreases.pdf"))`

`r rmdPageBreak()`

```{r 'analysis_agencyCounts_anomalousAgencies_dispositions', echo = FALSE}
if (!file.exists(txtFile) || !all(file.exists(imgFiles))) {
  percentages <- apply(dispoTable, c(1L, 2L), function(x) x["crt"] / sum(x))
  
  ratios <- apply(percentages[agenciesToKeep,], 1L, function(row) evalx(qlogis(tail(row, -1L)) - qlogis(head(row, -1L)), x[is.finite(x)]))
  increases <- sapply(ratios, function(x) if (length(x) > 0L) max(x) else 0)
  decreases <- sapply(ratios, function(x) if (length(x) > 0L) min(x) else 1)
  rm(ratios)
  
  ratioOrder <- order(increases, decreasing = TRUE)
  
  pdf(imgFiles[3L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, percentages[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         ylim = c(0, 1),
         main = agencyName, cex.main = 0.7,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByAgency[agenciesToKeep][ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  ratioOrder <- order(decreases)
  
  pdf(imgFiles[4L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, percentages[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         ylim = c(0, 1),
         main = agencyName, cex.main = 0.7,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByAgency[agenciesToKeep][ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  rm(increases, decreases)
  
  medians <- apply(percentages[agenciesToKeep,], 1L, function(row) median(row, na.rm = TRUE))
  
  ratioOrder <- order(medians, decreasing = TRUE)
  
  pdf(imgFiles[5L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, percentages[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         ylim = c(0, 1),
         main = agencyName, cex.main = 0.7,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByAgency[agenciesToKeep][ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  ratioOrder <- order(medians)
  
  pdf(imgFiles[6L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    agency <- subset(arrestingAgencies, code == levels(obts$arresting_agency)[agenciesToKeep][ratioOrder[i]])
    agencyName <- if (nrow(agency) > 0L) paste0(agency$agency, "\n", agency$type) else "Unknown"
    plot(years, percentages[which(agenciesToKeep)[ratioOrder[i]],], type = "l",
         ylim = c(0, 1),
         main = agencyName, cex.main = 0.7,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByAgency[agenciesToKeep][ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  rm(i, plotRegion, agency, agencyName, ratioOrder, medians)
}
```

#### Largest Increases - Agency Court Disposition Percentage

We perform a similar analysis, looking just at the court disposition percentage.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_dispositionIncreases.pdf"))`

`r rmdPageBreak()`

#### Largest Decreases - Agency Court Disposition Percentage

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_dispositionDecreases.pdf"))`

`r rmdPageBreak()`

#### Largest Agency Court Disposition Percentages

Finally, we order the agencies based on the median of their court disposition sentences across all years.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_dispositionMaxes.pdf"))`"

`r rmdPageBreak()`

#### Smallest Agency Court Disposition Percentages

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousAgencies_dispositionMins.pdf"))`"

`r rmdPageBreak()`

```{r echo = FALSE}
if (!file.exists(txtFile) || !all(file.exists(imgFiles)))
  rm(obts.sub, dispoTable, totalsByAgency, agenciesToKeep, years)
rm(txtFile, imgFiles)
```

## County Arrests Over Time

```{r 'analysis_agencyCounts_countyTotals', echo = FALSE}
txtFile <- file.path("..", txtPath, "analysis_agencyCounts_countyTotals.txt")
imgFile <- file.path("..", imgPath, "analysis_agencyCounts_countyTotals.pdf")
if (!file.exists(txtFile) || !file.exists(imgFile)) {
  loadData(mainTable, "..")
  
  county <- as.factor(substr(as.character(obts$arresting_agency), 1L, 2L))
  
  numCounties <- nlevels(county)
  
  writeLines(as.character(numCounties),
             txtFile)
  
  countyTable <- table(county)
  
  pdf(imgFile, 4.5, 4.5)
  par(mfrow = c(1L, 2L))
  hist(countyTable,
       breaks = 25L, freq = TRUE, xlab = "Num Arrests",
       main = "County Histogram")
  dev.off()
  
  rm(county, countyTable)
} else {
  numCounties <- as.integer(readLines(txtFile))
}
rm(txtFile, imgFile)
```

Due to natural circumstances, the arrests for some agencies can be moved to neighboring ones so that the variability shown above might be mitigated by looking at the county level. In contrast to the large number of agencies, only `r numCounties` counties are represented in the data.

```{r echo = FALSE}
rm(numCounties)
```

```{r 'analysis_agencyCounts_byCounty', echo = FALSE}
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_countyCounts.pdf",
                                       "analysis_agencyCounts_countyCourtDispositions.pdf"))
if (!all(file.exists(imgFiles))) {
  loadData(mainTable, "..")
  
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  courtDispositions <- c("acquitted", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                         "diversion_dismissed", "not_guilty_insane")
  releasedDispositions <- c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                            "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified")
  declinedDispositions <- c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                            "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                            "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                            "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable")
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                list(courtDispositions, releasedDispositions, declinedDispositions),
                c("crt", "lea", "da"))
  rm(courtDispositions, releasedDispositions, declinedDispositions)
  
  obts.sub$county <- as.factor(substr(as.character(obts.sub$arresting_agency), 1L, 2L))
  obts.sub <- obts.sub[,c("county", "court_event_year", "disposition")]
  
  years <- sort(unique(obts.sub$court_event_year))
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  
  totalsByCounty <- apply(totals, 1L, sum)
  percentages <- apply(dispoTable, c(1L, 2L), function(x) x["crt"]) / totals
  
  #loadData("judicialDistricts", "..")
  #countyNameMap <- unique(judicialDistricts[,c("county_code", "county")])
  
  
  #lowCounties <- which(apply(percentages, 1L, function(row) any(row <= 0.4, na.rm = TRUE)))
  #lowCountyInfo <- data.frame(code = names(lowCounties), name = countyNameMap$county[match(names(lowCounties), countyNameMap$county_code)])
  #lowCountyInfo$years <- sapply(lowCounties, function(index) {
  #  paste0(as.character(years[which(percentages[index,] <= 0.4)]), collapse = ", ")
  #})
  #lowCountyInfo$ave_yearly_count <- round(totalsByCounty[lowCounties] / length(years), 0)
  #lowCountyInfo <- lowCountyInfo[order(lowCountyInfo$ave_yearly_count, decreasing = TRUE),]
  #lowCountyInfo <- rmdFormat(lowCountyInfo, maxRows = NA)
  #rm(countyNameMap, lowCounties)
  
  cutoffs <- c(1000000L, 250000L, 100000L, 25000L, 10000L, 1L)
  
  indexSets <- list()
  indexSets[[1L]] <- unname(which(totalsByCounty > cutoffs[1L]))
  for (i in seq.int(2L, length(cutoffs))) indexSets[[i]] <- unname(which(cutoffs[i] < totalsByCounty & totalsByCounty <= cutoffs[i - 1L]))
  
  rm(totalsByCounty, cutoffs)
  
  pdf(imgFiles[1L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, max(totals[indices,])),
         xlab = "year", ylab = "count", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, totals[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
    plotRegion <- par()$usr
    xVal <- 0.975 * (plotRegion[2L] - plotRegion[1L]) + plotRegion[1L]
    yVal <- 0.975 * (plotRegion[4L] - plotRegion[3L]) + plotRegion[3L]
    text(xVal, yVal, paste0("n = ", length(indices)), adj = c(1, 0.75))
  }
  dev.off()
  
  pdf(imgFiles[2L], 6, 6)
  par(mfrow = getGridDim(8.5 / 11, length(indexSets)), mar = c(2.2, 2.2, 0.1, 0.1))
  
  for (i in seq_along(indexSets)) {
    indices <- indexSets[[i]]
    plot(NULL, type = "n", xlim = range(years), ylim = c(0, 1),
         xlab = "year", ylab = "% crt", xaxs = "i")
    for (j in seq_along(indices)) 
      lines(years, percentages[indices[j],], lwd = 1, col = rgb(0, 0, 0, 0.4))
  }
  dev.off()

  rm(i, j, indices, xVal, yVal, plotRegion)
  
  rm(years, indexSets, dispoTable, totals, percentages, obts.sub)
}
rm(imgFiles)
```

### County Totals

Here we replicate the agency-level distribution display, but this time with arbitrarily chosen cutoffs.

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_countyCounts.pdf"))`

`r rmdPageBreak()`

### County Court Disposition Percentages

Sharp dips in the arrests associated with a court disposition are likely due to reporting problems at the DA or court level. 

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_countyCourtDispositions.pdf"))`

`r rmdPageBreak()`

### Counties with Large Jumps

```{r 'analysis_agencyCounts_anomalousCounties', echo = FALSE}
imgFiles <- file.path("..", imgPath, c("analysis_agencyCounts_anomalousCounties_numArrestIncreases.pdf",
                                       "analysis_agencyCounts_anomalousCounties_numArrestDecreases.pdf",
                                       "analysis_agencyCounts_anomalousCounties_dispositionIncreases.pdf",
                                       "analysis_agencyCounts_anomalousCounties_dispositionDecreases.pdf",
                                       "analysis_agencyCounts_anomalousCounties_dispositionMaxes.pdf",
                                       "analysis_agencyCounts_anomalousCounties_dispositionMins.pdf"))
if (!all(file.exists(imgFiles))) {
  loadData(mainTable, "..")
  
  obts.sub <- obts[,c("arresting_agency", "court_event_year")]
  obts.sub$disposition <-
    as.factor(with(obts,
                   ifelse(!is.na(court_disposition_type), as.character(court_disposition_type), as.character(arrest_disposition_type))))

  courtDispositions <- c("acquitted", "certified_to_juvenile_court", "convicted", "defendant_deceased", "dismissed",
                         "diversion_dismissed", "not_guilty_insane")
  releasedDispositions <- c("lea_rel_arrestee_exonerated", "lea_rel_complaintant_refuses", "lea_rel_further_investigation",
                            "lea_rel_inadmissable_search", "lea_rel_lack_of_corpus", "lea_rel_other", "lea_rel_unspecified")
  declinedDispositions <- c("da_dec_combined_cases", "da_dec_complaintant_refuses", "da_dec_deferred_recovation_of_parole",
                            "da_dec_inadmissable_search", "da_dec_interest_of_justice", "da_dec_lack_of_cause",
                            "da_dec_lack_of_corpus", "da_dec_other", "da_dec_probation_revoked_1203_2PC",
                            "da_dec_prosecution_prefiling_deferred", "da_dec_witness_unavailable")
  obts.sub$disposition <-
    remapFactor(obts.sub$disposition,
                list(courtDispositions, releasedDispositions, declinedDispositions),
                c("crt", "lea", "da"))
  rm(courtDispositions, releasedDispositions, declinedDispositions)
  
  obts.sub$county <- as.factor(substr(as.character(obts.sub$arresting_agency), 1L, 2L))
  obts.sub <- obts.sub[,c("county", "court_event_year", "disposition")]
  
  dispoTable <- table(obts.sub)
  totals <- apply(dispoTable, c(1L, 2L), sum)
  totalsByCounty <- apply(totals, 1L, sum)
  
  loadData("judicialDistricts", "..")
  
  years <- sort(unique(obts.sub$court_event_year))
}
```

```{r 'analysis_agencyCounts_anomalousCounties_numArrests', echo = FALSE}
if (!all(file.exists(imgFiles))) {
  ratios <- apply(totals, 1L, function(row) evalx(tail(row, -1L) / head(row, -1L), x[is.finite(x) & x > 0]))
  increases <- sapply(ratios, function(x) if (length(x) > 0L) max(x) else 0)
  decreases <- sapply(ratios, function(x) if (length(x) > 0L) min(x) else 1)
  rm(ratios)
  
  ratioOrder <- order(increases, decreasing = TRUE)
  
  pdf(imgFiles[1L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, totals[ratioOrder[i],], type = "l",
         main = countyName,
         ylab = "num arrests", xlab = "year")
  }
  dev.off()
  
  ratioOrder <- order(decreases)
  
  pdf(imgFiles[2L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, totals[ratioOrder[i],], type = "l",
         main = countyName,
         ylab = "num arrests", xlab = "year")
  }
  dev.off()
  
  rm(i, county, countyName, ratioOrder, increases, decreases, totals)
}
```

#### Largest Increases - County Number of Arrests

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_numArrestIncreases.pdf"))`

`r rmdPageBreak()`

#### Largest Decreases - County Number of Arrests

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_numArrestDecreases.pdf"))`

`r rmdPageBreak()`

```{r 'analysis_agencyCounts_anomalousCounties_dispositions', echo = FALSE}
if (!all(file.exists(imgFiles))) {
  percentages <- apply(dispoTable, c(1L, 2L), function(x) x["crt"] / sum(x))
  
  ratios <- apply(percentages, 1L, function(row) evalx(qlogis(tail(row, -1L)) - qlogis(head(row, -1L)), x[is.finite(x)]))
  increases <- sapply(ratios, function(x) if (length(x) > 0L) max(x) else 0)
  decreases <- sapply(ratios, function(x) if (length(x) > 0L) min(x) else 1)
  rm(ratios)
  
  ratioOrder <- order(increases, decreasing = TRUE)
  
  pdf(imgFiles[3L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, percentages[ratioOrder[i],], type = "l",
         ylim = c(0, 1),
         main = countyName,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByCounty[ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  ratioOrder <- order(decreases)
  
  pdf(imgFiles[4L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, percentages[ratioOrder[i],], type = "l",
         ylim = c(0, 1),
         main = countyName,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByCounty[ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  rm(increases, decreases)
  
  medians <- apply(percentages, 1L, function(row) median(row, na.rm = TRUE))
  
  ratioOrder <- order(medians, decreasing = TRUE)
  
  pdf(imgFiles[5L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, percentages[ratioOrder[i],], type = "l",
         ylim = c(0, 1),
         main = countyName,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByCounty[ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  ratioOrder <- order(medians)
  
  pdf(imgFiles[6L], 6L, 6L)
  par(mfrow = c(3L, 3L))
  for (i in seq_len(9L)) {
    county <- subset(countyNameMap, county_code == levels(obts.sub$county)[ratioOrder[i]])
    countyName <- if (nrow(county) > 0L) paste0(county$county) else "Unknown"
    plot(years, percentages[ratioOrder[i],], type = "l",
         ylim = c(0, 1),
         main = countyName,
         ylab = "% crt", xlab = "year")
    plotRegion <- par("usr")
    mtext(paste0("n = ", totalsByCounty[ratioOrder[i]]),
          at = c(plotRegion[2L], plotRegion[4L]), adj = c(1, 0.5),
          cex = 0.6)
  }
  dev.off()
  
  rm(i, plotRegion, county, countyName, ratioOrder, medians, percentages)
}
```

#### Largest Increases - County Court Disposition Percentage

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_dispositionIncreases.pdf"))`

`r rmdPageBreak()`

#### Largest Decreases - County Court Disposition Percentage

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_dispositionDecreases.pdf"))`

`r rmdPageBreak()`

#### Largest County Court Disposition Percentages

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_dispositionMaxes.pdf"))`"

`r rmdPageBreak()`

#### Smallest County Court Disposition Percentages

`r rmdImageInline(file.path(imgPath, "analysis_agencyCounts_anomalousCounties_dispositionMins.pdf"))`"

`r rmdPageBreak()`

```{r echo = FALSE}
if (!all(file.exists(imgFiles)))
  rm(obts.sub, dispoTable, totalsByCounty, years)
rm(imgFiles)
```