# Table Joins

Some entries in the OBTS do not have matching values in the related tables, specifically the judicial districts, arresting agencies, and CJIS codes.

## Judicial Districts

```{r 'analysis_joinFailures_judicialDistricts', echo = FALSE, results = "asis"}
txtFile <- file.path("..", txtPath, "analysis_joinFailures_judicialDistricts.txt")
if (!file.exists(txtFile)) {
  loadData("judicialDistricts", "..")
  loadData(mainTable, "..")
  districtCodes <- with(judicialDistricts, paste0(county_code, ifelse(is.na(agency_code), "****", agency_code)))
  unmatchedRows <- obts$court_judicial_district %not_in% districtCodes
  
  unmatchedSet <- subset(obts, unmatchedRows, c("court_judicial_district", "court_event_year"))
  unmatchedSet$court_judicial_district <- droplevels(unmatchedSet$court_judicial_district)
  
  numUnmatched <- sum(unmatchedRows)
  unmatchedByJurisdiction <- rmdFormat(sort(table(unmatchedSet$court_judicial_district), decreasing = TRUE))
  unmatchedByYear <- rmdFormat(table(unmatchedSet$court_event_year), maxRows = NA)
  
  writeLines(c(as.character(numUnmatched),
               unmatchedByJurisdiction,
               unmatchedByYear),
             txtFile)
  rm(districtCodes, unmatchedRows, unmatchedSet)
} else {
  lines <- readLines(txtFile)
  numUnmatched <- lines[1L]
  unmatchedByJurisdiction <- lines[seq.int(2L, 54L)]
  unmatchedByYear <- tail(lines, -54L)
  rm(lines)
}
rm(txtFile)
```

A total of `r numUnmatched` rows have listed judicial districts that cannot be matched.

#### Unmatched by Jurisdiction Tabulation

`r paste0(unmatchedByJurisdiction, collapse = "\n")`

#### Unmatched by Court Event Year Tabulation

`r paste0(unmatchedByYear, collapse = "\n")`

```{r, echo = FALSE}
rm(numUnmatched, unmatchedByJurisdiction, unmatchedByYear)
```

## Arresting Agencies

```{r 'analysis_joinFailures_arrestingAgencies', echo = FALSE, results = "asis"}
txtFile <- file.path("..", txtPath, "analysis_joinFailures_arrestingAgencies.txt")
if (!file.exists(txtFile)) {
  loadData("arrestingAgencies", "..")
  loadData(mainTable, "..")
  unmatchedRows <- obts$arresting_agency %not_in% arrestingAgencies$code
  
  unmatchedSet <- subset(obts, unmatchedRows, c("arresting_agency", "court_event_year"))
  unmatchedSet$arresting_agency <- droplevels(unmatchedSet$arresting_agency)
  
  numUnmatched <- sum(unmatchedRows)
  unmatchedByAgency <- rmdFormat(sort(table(unmatchedSet$arresting_agency), decreasing = TRUE))
  unmatchedByYear <- rmdFormat(table(unmatchedSet$court_event_year), maxRows = NA)
  
  writeLines(c(as.character(numUnmatched),
               unmatchedByAgency,
               unmatchedByYear),
             txtFile)
  rm(unmatchedRows, unmatchedSet)
} else {
  lines <- readLines(txtFile)
  numUnmatched <- lines[1L]
  unmatchedByAgency <- lines[seq.int(2L, 54L)]
  unmatchedByYear <- tail(lines, -54L)
  rm(lines)
}
rm(txtFile)
```

A total of `r numUnmatched` rows have listed arresting agencies that cannot be matched.

#### Unmatched by Agency Tabulation

`r paste0(unmatchedByAgency, collapse = "\n")`

#### Unmatched by Court Event Year Tabulation

`r paste0(unmatchedByYear, collapse = "\n")`

```{r, echo = FALSE}
rm(numUnmatched, unmatchedByAgency, unmatchedByYear)
```

## Offense Codes

```{r echo = FALSE}
## helper function

## merges together two univariate tables into a dataframe with the first column being
## the names of the values and the second being the counts
joinTables <- function(t1, t2, names = NULL) {
  df <- data.frame(name = union(base::names(t1), base::names(t2)), t1 = 0L, t2 = 0L)
  df[match(base::names(t1), df$name), "t1"] <- unname(t1)
  df[match(base::names(t2), df$name), "t2"] <- unname(t2)
  
  df <- df[order(df$t1 + df$t2, decreasing = TRUE),]
  if (!is.null(names)) base::names(df) <- names
  df
}
```

```{r 'analysis_joinFailures_offenseCodes', echo = FALSE, results = "asis"}
txtFile <- file.path("..", txtPath, "analysis_joinFailures_offenseCodes.txt")
if (!file.exists(txtFile)) {
  loadData("cjisCodes", "..")
  loadData(mainTable, "..")
  
  while (sum(shortCodes <- nchar(levels(obts$arrest_offense)) < 5L) > 0L)
    levels(obts$arrest_offense)[shortCodes] <- paste0("0", levels(obts$arrest_offense)[shortCodes])
  
  while (sum(shortCodes <- nchar(levels(obts$court_disposition_offense)) < 5L) > 0L)
    levels(obts$court_disposition_offense)[shortCodes] <- paste0("0", levels(obts$court_disposition_offense)[shortCodes])
  rm(shortCodes)
  
  unmatchedArrestOffenseRows <- obts$arrest_offense %not_in% cjisCodes$cjis_code & !is.na(obts$arrest_offense)
  unmatchedArrestOffenses <- droplevels(obts$arrest_offense[unmatchedArrestOffenseRows])
  
  unmatchedCourtDispositionOffenseRows <- obts$court_disposition_offense %not_in% cjisCodes$cjis_code & !is.na(obts$court_disposition_offense)
  unmatchedCourtDispositionOffenses <- droplevels(obts$court_disposition_offense[unmatchedCourtDispositionOffenseRows])
  
  numUnmatchedArrestOffenses <- length(unmatchedArrestOffenses)
  numUnmatchedCourtDispositionOffenses <- length(unmatchedCourtDispositionOffenses)
  
  unmatchedArrestOffenseTable <- as.data.frame(table(unmatchedArrestOffenses))
  unmatchedCourtDispositionOffenseTable <- as.data.frame(table(unmatchedCourtDispositionOffenses))
  
  unmatchedOffensesTable <- joinTables(unmatchedArrestOffenseTable, unmatchedCourtDispositionOffenseTable,
                                       c("code", "arrest_offense", "court_disposition_offense"))
  unmatchedByOffense <-
    rmdFormat(unmatchedOffensesTable[order(unmatchedOffensesTable[[2L]] + unmatchedOffensesTable[[3L]], decreasing = TRUE),])
  
  unmatchedArrestOffenseTable <- table(obts$court_event_year[unmatchedArrestOffenseRows])
  unmatchedCourtDispositionOffenseTable <- table(obts$court_event_year[unmatchedCourtDispositionOffenseRows])
  
  unmatchedYearsTable <- joinTables(unmatchedArrestOffenseTable, unmatchedCourtDispositionOffenseTable,
                                    c("court_event_year", "arrest_offense", "court_disposition_offense"))
  unmatchedYearsTable <- unmatchedYearsTable[order(unmatchedYearsTable$court_event_year),]
  
  unmatchedByYear <- rmdFormat(unmatchedYearsTable[order(unmatchedYearsTable$court_event_year),], maxRows = NA)
  
  
  
  writeLines(c(as.character(numUnmatchedArrestOffenses),
               as.character(numUnmatchedCourtDispositionOffenses),
               unmatchedByOffense,
               unmatchedByYear),
             txtFile)
  
  rm(unmatchedArrestOffenseRows, unmatchedArrestOffenses, unmatchedCourtDispositionOffenseRows,
     unmatchedCourtDispositionOffenses, unmatchedOffensesTable, unmatchedYearsTable,
     unmatchedArrestOffenseTable, unmatchedCourtDispositionOffenseTable)
} else {
  lines <- readLines(txtFile)
  numUnmatchedArrestOffenses <- as.integer(lines[1L])
  numUnmatchedCourtDispositionOffenses <- as.integer(lines[2L])
  unmatchedByOffense <- lines[seq.int(3L, 55L)]
  unmatchedByYear <- tail(lines, -55L)
  rm(lines)
}
rm(txtFile)
```

A total of `r numUnmatchedArrestOffenses` rows have arrest offense codes that cannot be matched, while `r numUnmatchedCourtDispositionOffenses` rows have court disposition offense codes that cannot be matched.

#### Unmatched by Offense Code Tabulation

`r paste0(unmatchedByOffense, collapse = "\n")`

#### Unmatched by Court Event Year Tabulation

`r paste0(unmatchedByYear, collapse = "\n")`

```{r, echo = FALSE}
rm(numUnmatchedArrestOffenses, numUnmatchedCourtDispositionOffenses, unmatchedByOffense, unmatchedByYear)
```

## Qualifier Codes

```{r 'analysis_joinFailures_qualifierCodes', echo = FALSE, results = "asis"}
txtFile <- file.path("..", txtPath, "analysis_joinFailures_qualifierCodes.txt")
if (!file.exists(txtFile)) {
  loadData("qualifierCodes", "..")
  loadData(mainTable, "..")
  
  arrestQualifierIndices <- t(sapply(as.character(obts$arrest_qualifier), function(qualifier) {
    qualifiers <- substring(qualifier, seq.int(1L, 13L, 2L), seq.int(2L, 14L, 2L))
    indices <- match(qualifiers, qualifierCodes$qualifier)
    unmatchedRows <- is.na(indices) & qualifiers != "  "
    if (any(unmatchedRows)) indices[unmatchedRows] <- 0L
    indices
  }))
  
  unmatchedArrestQualifierRows <- apply(arrestQualifierIndices, 1L, function(row) any(row == 0L, na.rm = TRUE))
  
  unmatchedArrestQualifiers <- lapply(which(unmatchedArrestQualifierRows), function(i) {
    qualifierIndices <- arrestQualifierIndices[i,] == 0L & !is.na(arrestQualifierIndices[i,])
    startIndices <- seq.int(1L, 13L, 2L)[qualifierIndices]
    endIndices   <- seq.int(2L, 14L, 2L)[qualifierIndices]
    substring(obts$arrest_qualifier[i], startIndices, endIndices)
  })
  
  unmatchedArrestQualifiersTable <- table(unlist(unmatchedArrestQualifiers))
  
  
  courtQualifierIndices <- t(sapply(as.character(obts$court_qualifier), function(qualifier) {
    qualifiers <- substring(qualifier, seq.int(1L, 13L, 2L), seq.int(2L, 14L, 2L))
    indices <- match(qualifiers, qualifierCodes$qualifier)
    unmatchedRows <- is.na(indices) & qualifiers != "  "
    if (any(unmatchedRows)) indices[unmatchedRows] <- 0L
    indices
  }))
  
  unmatchedCourtQualifierRows <- apply(courtQualifierIndices, 1L, function(row) any(row == 0L, na.rm = TRUE))
  
  unmatchedCourtQualifiers <- lapply(which(unmatchedCourtQualifierRows), function(i) {
    qualifierIndices <- courtQualifierIndices[i,] == 0L & !is.na(courtQualifierIndices[i,])
    startIndices <- seq.int(1L, 13L, 2L)[qualifierIndices]
    endIndices   <- seq.int(2L, 14L, 2L)[qualifierIndices]
    substring(obts$court_qualifier[i], startIndices, endIndices)
  })
  
  unmatchedCourtQualifiersTable <- table(unlist(unmatchedCourtQualifiers))
  
  unmatchedByQualifierTable <- joinTables(unmatchedArrestQualifiersTable, unmatchedCourtQualifiersTable,
                                          c("code", "arrest_qualifier", "court_qualifier"))
  
  unmatchedByYearTable <- joinTables(table(obts$court_event_year[unmatchedArrestQualifierRows]),
                                     table(obts$court_event_year[unmatchedCourtQualifierRows]),
                                     c("court_event_year", "arrest_qualifier", "court_qualifier"))
  unmatchedByYearTable <- unmatchedByYearTable[order(unmatchedByYearTable$court_event_year),]
  
  numUnmatchedArrestQualifiers <- sum(unmatchedArrestQualifierRows)
  numUnmatchedCourtQualifiers  <- sum(unmatchedCourtQualifierRows)
  unmatchedByQualifier <- rmdFormat(unmatchedByQualifierTable)
  unmatchedByYear      <- rmdFormat(unmatchedByYearTable, maxRows = NA)
  
  writeLines(c(as.character(numUnmatchedArrestQualifiers),
               as.character(numUnmatchedCourtQualifiers),
               unmatchedByQualifier,
               unmatchedByYear),
             txtFile)
  
  rm(arrestQualifierIndices, unmatchedArrestQualifierRows, unmatchedArrestQualifiers, unmatchedArrestQualifiersTable,
     courtQualifierIndices, unmatchedCourtQualifierRows, unmatchedCourtQualifiers, unmatchedCourtQualifiersTable,
     unmatchedByQualifierTable, unmatchedByYearTable)
} else {
  lines <- readLines(txtFile)
  numUnmatchedArrestQualifiers <- as.integer(lines[1L])
  numUnmatchedCourtQualifiers <- as.integer(lines[2L])
  unmatchedByQualifier <- lines[seq.int(3L, 55L)]
  unmatchedByYear <- tail(lines, -55L)
  rm(lines)
}
rm(txtFile)
```

A total of `r numUnmatchedArrestQualifiers` rows have arrest qualifier codes that cannot be matched, while `r numUnmatchedCourtQualifiers` rows have court disposition offense codes that cannot be matched. Because rows can have multiple qualifier codes, the following tables do not necessarily add up to these numbers.

#### Unmatched by Offense Code Tabulation

`r paste0(unmatchedByQualifier, collapse = "\n")`

#### Unmatched by Court Event Year Tabulation

`r paste0(unmatchedByYear, collapse = "\n")`

```{r, echo = FALSE}
rm(joinTables)
```