---
title: "MACR Statistical Disclosure Analysis"
date: "November 30, 2016"
output: pdf_document
---

```{r 'data_prep', include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

commonSrcPath  <- file.path("..", "common", "src")
commonDataPath <- file.path("..", "common", "data")

jurisdictions <- read.csv(file.path(commonDataPath, "jurisdictions.csv"), stringsAsFactors = FALSE)
bcsOffenseCodes <- read.csv(file.path(commonDataPath, "bcs_offense_codes.csv"), stringsAsFactors = FALSE)

source(file.path(commonSrcPath, "knitr.R"))
source(file.path(commonSrcPath, "util.R"))
source(file.path(commonSrcPath, "jurisdictions.R"))
source(file.path(commonSrcPath, "bcsOffenseCodes.R"))

source(file.path("..", "anonymization", "src", "suppression.R"))

load(file.path("datasets", "macr", "macr_clean.Rdata"))
years <- sort(unique(macr$arrest_year))
numYears <- length(years)

## drop minors
macr.puf <- subset(macr, age > 17L)

## add county
countyJurisdictionMap <- as.integer(substr(levels(macr.puf$ncic_jurisdiction), 1L, 2L))
macr.puf$county <- countyJurisdictionMap[as.integer(macr.puf$ncic_jurisdiction)]
rm(countyJurisdictionMap)

## map specific Asian groups to "Asian or Pac Isl" and small, non-Asian groups to "Other"
asianCategoryNames <- c("Chinese", "Asian Indian", "Cambodian", "Guamanian", "Filipino", "Hawaiian", "Japanese", "Korean", "Laotian", "other Asian", "Pacific Islander", "Samoan", "Vietnamese")

countyRemapNames <- list(
  c("Alpine County", "Tuolumne County", "Calaveras County", "Amador County"),
  c("Colusa County", "Glenn County", "Yuba County", "Sutter County", "Lake County"),
  c("Sierra County", "Plumas County", "Lassen County", "Modoc County", "Nevada County"),
  c("Inyo County", "Mono County", "Mariposa County", "Madera County", "San Benito County"),
  c("Trinity County", "Del Norte County", "Siskiyou County", "Tehama County"))
countyNameMap <- unique(jurisdictions[,c("CntyCode", "County")])
colnames(countyNameMap) <- c("code", "name")

countyRemap <- lapply(countyRemapNames, function(names) countyNameMap$code[match(names, countyNameMap$name)])

countyNameMap$name <- gsub("(.*) County", "\\1", countyNameMap$name, perl = TRUE)

countyNameMap$code.new <- countyNameMap$code
countyNameMap$name.new <- countyNameMap$name

## replaces each code/county name with min matching code and collapsed name
for (i in seq_along(countyRemap)) {
  remapRows <- match(countyRemap[[i]], countyNameMap$code.new)
  countyNameMap$code.new[remapRows[-1L]] <- countyRemap[[i]][1L]
  countyNameMap$name.new[remapRows] <- paste0(countyNameMap$name.new[remapRows], collapse = "/")
}

## reorders codes starting from 1
countyNameMap$code.new <- match(countyNameMap$code.new, sort(unique(countyNameMap$code.new)))
```

## Aggregating/Recoding/Binning

We begin by dropping the direct identifiers and redefining the quasi identifiers as coarse grouping variables.

```{r 'recode', echo = FALSE}
macr.puf <- macr.puf[,c("arrest_year", "county", "gender", "race", "age", "summary_offense_level", "bcs_summary_offense_code", "disposition")]

## rename columns to be less obscure
colnames(macr.puf) <- c("year", "county", "gender", "race", "age", "offense_level", "offense_code", "disposition")

## if we use the sdc code for this, it recomputes risk after every stage, making it take an incredibly long time

## county is no longer a factor, so we re-code by hand
macr.puf$county <- countyNameMap$code.new[match(macr.puf$county, countyNameMap$code)]
macr.puf$county <- as.factor(macr.puf$county)

macr.puf$race <- remapFactor(macr.puf$race, list(asianCategoryNames, "American Indian"), c("Asian or Pac Isl", "other"))
  
macr.puf$age <- cut(macr.puf$age, c(17, 24, 32, 44, 150), labels = c("18-24", "25-32", "33-44", "45+"))

macr.puf$disposition <- remapFactor(macr.puf$disposition, c("felony complaint sought", "misdemeanor complaint sought"), "complaint sought")

for (i in seq_along(macr.puf))
  if (is.factor(macr.puf[[i]])) macr.puf[[i]] <- droplevels(macr.puf[[i]])

invisible(gc(FALSE))
```

## Protecting That Arrested

```{r 'that_arrested', echo = FALSE, results = "asis"}



subRows <- with(macr.puf, disposition %in% "released")
selectOffenseCodes <- with(bcsSummaryOffenseCodes, summary_offense_code[match(c("Prostitution", "Drunk"), summary_offense_type)])
releasedResult <- enforceKAnonymity(macr.puf[subRows,],
                                    across = "year",
                                    inQuasiIdentifiers = c("gender", "race", "age"),
                                    suppressing = "county",
                                    k = 5L,
                                    matchingCriteria = macr.puf$offense_code[subRows] %in% selectOffenseCodes)
macr.puf[subRows, "county"] <- releasedResult$suppressed

subRows <- with(macr.puf, disposition %in% "complaint sought" & offense_level %in% "felony")
felonyReferredResult <- enforceKAnonymity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          suppressing = "county",
                                          k = 5L)
macr.puf[subRows, "county"] <- felonyReferredResult$suppressed

subRows <- with(macr.puf, disposition %in% "released" & offense_level %in% "felony")
felonyReleasedResult <- enforceKAnonymity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          suppressing = "county",
                                          k = 5L)
macr.puf[subRows, "county"] <- felonyReleasedResult$suppressed

subRows <- with(macr.puf, disposition %not_in% "complaint sought" | offense_level %not_in% "felony")
notFelonyReferredResult <- enforceKAnonymity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          suppressing = "county",
                                          k = 5L)
macr.puf[subRows, "county"] <- notFelonyReferredResult$suppressed

totalPopulationResult <- enforceKAnonymity(macr.puf,
                                           across = "year",
                                           inQuasiIdentifiers = c("gender", "race", "age"),
                                           suppressing = "county",
                                           k = 5L)
macr.puf[,"county"] <- totalPopulationResult$suppressed

numSuppressed <- cbind(releasedResult$numSuppressed, felonyReferredResult$numSuppressed, notFelonyReferredResult$numSuppressed,
                       felonyReleasedResult$numSuppressed, totalPopulationResult$numSuppressed)
colnames(numSuppressed) <- c("k-rel #", "k-rel %", "k-fref #", "k-fref %", "k-!frf #", "k-!frf %",
                             "k-frel #", "k-frel %", "k-tot #", "k-tot %")

cat(rmdFormat(numSuppressed, digits = 2L), sep = "\n")
rm(numSuppressed, releasedResult, felonyReferredResult, notFelonyReferredResult, felonyReleasedResult, totalPopulationResult)
```

After this step, the number of rows with a suppressed county variable is `r sum(is.na(macr.puf$county))` / `r nrow(macr.puf)`, or `r round(100 * sum(is.na(macr.puf$county)) / nrow(macr.puf), 2)`%.

## Protecting Arrest Offense

```{r 'arrest_offense', echo = FALSE, results = "asis"}
subRows <- with(macr.puf, disposition %in% "complaint sought" & offense_level %in% "felony")
felonyReferredResult <- enforceLDiversity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          forSensitiveValues = "offense_code",
                                          suppressing = "county",
                                          l = 5L)
macr.puf[subRows, "county"] <- felonyReferredResult$suppressed

subRows <- with(macr.puf, disposition %not_in% "complaint sought" | offense_level %not_in% "felony")
notFelonyReferredResult <- enforceLDiversity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          forSensitiveValues = "offense_code",
                                          suppressing = "county",
                                          l = 5L)
macr.puf[subRows, "county"] <- notFelonyReferredResult$suppressed

subRows <- with(macr.puf, disposition %in% "released" & offense_level %in% "felony")
felonyReleasedResult <- enforceLDiversity(macr.puf[subRows,],
                                          across = "year",
                                          inQuasiIdentifiers = c("gender", "race", "age"),
                                          forSensitiveValues = "offense_code",
                                          suppressing = "county",
                                          l = 5L)
macr.puf[subRows, "county"] <- felonyReleasedResult$suppressed

totalPopulationResult <- enforceLDiversity(macr.puf,
                                           across = "year",
                                           inQuasiIdentifiers = c("gender", "race", "age"),
                                           forSensitiveValues = "offense_code",
                                           suppressing = "county",
                                           l = 5L)
macr.puf[,"county"] <- totalPopulationResult$suppressed

numSuppressed <- cbind(felonyReferredResult$numSuppressed, notFelonyReferredResult$numSuppressed,
                       felonyReleasedResult$numSuppressed, totalPopulationResult$numSuppressed)
colnames(numSuppressed) <- c("l-fref #", "l-fref %", "l-!frf #", "l-!frf %",
                             "l-frel #", "l-frel %", "l-tot #", "l-tot %")

cat(rmdFormat(numSuppressed, digits = 2L), sep = "\n")
rm(numSuppressed, felonyReferredResult, notFelonyReferredResult, felonyReleasedResult, totalPopulationResult)
```

After this step, the number of rows with a suppressed county variable is `r sum(is.na(macr.puf$county))` / `r nrow(macr.puf)`, or `r round(100 * sum(is.na(macr.puf$county)) / nrow(macr.puf), 2)`%.

## Result

```{r 'save_result', echo = FALSE}
saveResult <- TRUE

if (saveResult) {
  dir.create(file.path("datasets", "macr", "puf"), recursive = TRUE)
  
  counties <- countyNameMap[,c("code.new", "name.new")]
  colnames(counties) <- c("code", "name")
  counties <- counties[!duplicated(counties),]
  counties <- counties[order(counties$code),]
  
  offenseCodes <- bcsSummaryOffenseCodes
  colnames(offenseCodes) <- c("code", "type")
  
  macr.puf$county <- with(macr.puf, as.integer(levels(county))[county])
  macr.puf$offense_code <- with(macr.puf, as.integer(levels(offense_code))[offense_code])
  
  write.csv(counties, file = file.path("datasets", "macr", "puf", "counties.csv"), row.names = FALSE)
  write.csv(offenseCodes, file = file.path("datasets", "macr", "puf", "offenseCodes.csv"), row.names = FALSE)
  write.csv(macr.puf, file = file.path("datasets", "macr", "puf", "macr.csv"), row.names = FALSE)
}
```

Total number of rows with a suppressed county variable: `r sum(is.na(macr.puf$county))` / `r nrow(macr.puf)` = `r sum(is.na(macr.puf$county)) / nrow(macr.puf)`.
