# Booking Rates

This section examines the reliability of booking data from the MACR database (the status_type field). We find that booking data appears to be extremely noisy and generally biased upward. In other words, MACR data about whether an individual was cited or booked is unreliable; an implausibly high percentage of arrests are recorded as booked. We conclude that researchers who decide they want to use these data must be extremely cautious in assessing the data for themselves and in deciding how to use them. We provide a variable loosely describing the quality of booking data by department, but these quality measures are based on heavy assumptions and should not be the only basis for deciding which data to use.  
 
The first reliability analyses focus on the plausibility of booking rates given what we know from talking with law enforcement personnel. For instance, we consider whether a jurisdiction reports booking 100% of arrests or whether a jurisdiction tends to book juveniles for misdemeanors. In discussions with law enforcement personnel (officers, IT staff, department analysts), we learned that officers typically cite and release for misdemeanors and book for felonies. Individuals we spoke with said they would be skeptical if a department claimed it booked the majority of misdemeanors (only those which pose potential for immediate harm, such as domestic violence, are candidates for booking). Booking someone takes at least 1.5-2 hours, potentially diverting an officer from responding to calls, and requiring a great deal of tedious paperwork. Individuals who are on probation or parole who are arrested for a misdemeanor must be booked (roughly 22% of individuals arrested in CA are on probation or parole). In addition, individuals who have a warrant out must be booked, and the MACR will show the new crime for which they were arrested. If an individual was stopped, but not arrested until the officer discovered the outstanding warrant, the warrant offense code will be shown on the MACR. 
  
The second portion of the reliability analyses are based on a probabilistic matching of reported bookings to actual bookings. The reported bookings come from the MACR, while the actual bookings are drawn from the Automated Criminal History System (ACHS). Records in ACHS are created when someone is fingerprinted during jail intake (i.e., booking). Due to data quality issues and because the matching was probabilistic, there are fewer matched bookings in our analyses than there would be if the matching process matching were perfect. This has the overall effect of reducing the percentage of reports of bookings that appear accurate. Only about 12% of reported bookings could be matched to an actual booking. The percentage match by department is normally distributed around this figure. In other words, for the average department, we can only confirm that roughly 12% of reported bookings had a corresponding record in a database of actual bookings. We do not know if 12% therefore reflects "accurate" reporting or whether there are systematic biases that have reduced the match percentage beyond artefacts of the data linkage process. Both systematic biases and data linkage problems may produce the figure.  

One way to examine the plausibilty of booking data by department is to consider the severity of the offense. We expected that arrests for more severe crimes would have a higher match percentage because they were more likely to have actually been booked. Arrests for less serious crimes, that were in fact cited, would have a lower percentage match. Surprisingly, the booking match percentage did not vary much by the severity of offense. The booking match percentage only declines slightly from homicide to misdemeanors. One unexplored potential explanation is that there are missing data about arrests for less serious crimes (perhaps due to automatic entry of bookings into the agency's record-management system). In other words, we may not know the real denominator of total arrests for less serious crimes. This would affect studies that examine the propensity to book arrested individuals across department, crime, or demographic group.

```{r 'analysis_bookingRates_cache', echo = FALSE}
txtFiles <- file.path("..", txtPath, c("analysis_bookingRates_juvenileMisdemeanors_1.txt",
                                       "analysis_bookingRates_juvenileMisdemeanors_2.txt",
                                       "analysis_bookingRates_adultFelonies.txt",
                                       "analysis_bookingRates_remaining.txt",
                                       "analysis_bookingRates_felonyTable.csv",
                                       "analysis_bookingRates_misdemeanorTable.csv"))
imgFiles <- file.path("..", imgPath, c("analysis_bookingRates_juvenileMisdemeanors_1.pdf",
                                       "analysis_bookingRates_juvenileMisdemeanors_2.pdf",
                                       "analysis_bookingRates_juvenileMisdemeanors_3.pdf",
                                       "analysis_bookingRates_adultFelonies_1.pdf",
                                       "analysis_bookingRates_adultFelonies_2.pdf",
                                       "analysis_bookingRates_bookingProportions.pdf",
                                       "analysis_bookingRates_logStat.pdf",
                                       "analysis_bookingRates_bookedMatchedDensity.pdf",
                                       "analysis_bookingRates_actualBookedHist.pdf",
                                       "analysis_bookingRates_repBookedVDiff.pdf",
                                       "analysis_bookingRates_remaining_1.pdf",
                                       "analysis_bookingRates_remaining_2.pdf"))

runAnalysis <- !all(file.exists(txtFiles)) || !all(file.exists(imgFiles))
if (runAnalysis) {
  macr <- loadData(path = "..")
  bcsOffenseCodes <- loadData("bcs_offense_codes", path = "..")
  
  macr.sub <- subset(macr, arrest_year >= 2005,
                     c("ncic_jurisdiction", "bcs_summary_offense_code", "status_type",
                       "offense_level",     "arrest_year",              "age",
                       "race_or_ethnicity"))
  numYears <- length(unique(macr.sub$arrest_year))
  
  macr.sub$ncic_jurisdiction <- droplevels(macr.sub$ncic_jurisdiction)
  
  source(file.path("src", "analysis_bookingRates_plot.R"))
}
```

## Eliminating Jurisdictions

This section uses the MACR file to aggregate and display booking rates for felonies and misdemeanors by first removing jurisdictions with anomalous booking rates.

### Sample Size

The first cut we make is to drop jurisdictions that make a small number of arrests overall. By considering the cumulative distribution of the number of arrests, we can set cut points at arbitrary percentages. To start we include only those jurisdictions responsible for 95% of all arrests.

```{r 'analysis_bookingRates_sample_size', echo = FALSE}
if (runAnalysis) {
  numArrestsByJurisdiction <- table(macr.sub$ncic_jurisdiction)
  
  jurisdictionOrder <- order(unname(numArrestsByJurisdiction))
  numArrestsByJurisdiction.pcnt <-
    cumsum(numArrestsByJurisdiction[jurisdictionOrder]) / sum(numArrestsByJurisdiction)
  
  ## keep top 95%
  largeJurisdictionIndices <-
    jurisdictionOrder[seq.int(which.min(numArrestsByJurisdiction.pcnt <= 0.05),
                              length(numArrestsByJurisdiction))]
  jurisdictionsToKeep <- names(numArrestsByJurisdiction)[largeJurisdictionIndices]
  
  macr.sub <- subset(macr.sub, ncic_jurisdiction %in% jurisdictionsToKeep)
  macr.sub$ncic_jurisdiction <- droplevels(macr.sub$ncic_jurisdiction)
  
  ## clean up environment a bit to make sure we don't have accidental name collisions later
  rm(numArrestsByJurisdiction, jurisdictionOrder, numArrestsByJurisdiction.pcnt,
     largeJurisdictionIndices, jurisdictionsToKeep)
  invisible(gc(FALSE))
}
```

### Abnormal Booking Rates

Our goal at this stage is to assess the reliability of jurisdictions in their reporting. We can look for jursidictions that are reporting abornomally high booking rates for less serious crimes or low booking rates for serious crimes. We first focus on booking rates of juveniles for less serious crimes. An analysis of booking rates for individual offense codes (not shown here) indicates a large amount of variability across years and jursidictions, so we instead consider measures that aggregate arrests *before* computing booking rates. That is, we define rates such as:

$$
\text{mostly-harmless-booking-rate} = \dfrac{\text{num juveniles booked for selected codes}}{\text{total num juveniles arrested for selected codes}}
$$

```{r 'analysis_bookingRates_functions', echo = FALSE}
if (runAnalysis) {
  computeBookingRate <- function(x) {
    if (sum(x) == 0L) return(NA)
    unname(x["booked"] / sum(x))
  }
  
  getJurisdictionalBookingRatesForOffenseCodes <- function(macr, offenseCodes) {
    ## drop unnecessary columns
    macr.sub <- subset(macr, bcs_summary_offense_code %in% offenseCodes,
                       c("ncic_jurisdiction", "bcs_summary_offense_code", "status_type"))
    
    tab <- table(macr.sub)
    apply(tab, c(1L, 2L), computeBookingRate)
  }
  getJurisdictionalBookingRatesForOffenseCodesByYear <- function(macr, offenseCodes) {
    macr.sub <- subset(macr, bcs_summary_offense_code %in% offenseCodes,
                       c("ncic_jurisdiction", "bcs_summary_offense_code", "arrest_year", "status_type"))
    
    tab <- table(macr.sub)
    apply(tab, c(1L, 2L, 3L), computeBookingRate)
  }
  ## aggregates by summing all in set
  getJurisdictionalBookingRatesForOffenseCodeSet <- function(macr, offenseCodes) {
    macr.sub <- subset(macr, bcs_summary_offense_code %in% offenseCodes,
                       c("ncic_jurisdiction", "bcs_summary_offense_code", "status_type"))
    
    tab <- table(macr.sub)
    tab <- apply(tab, c(1L, 3L), sum, na.rm = TRUE)
    
    apply(tab, 1L, computeBookingRate)
  }
  getJurisdictionalBookingRatesForOffenseCodeSetByYear <- function(macr, offenseCodes) {
    macr.sub <- subset(macr, bcs_summary_offense_code %in% offenseCodes,
                       c("ncic_jurisdiction", "bcs_summary_offense_code", "arrest_year", "status_type"))
    
    tab <- table(macr.sub)
    tab <- apply(tab, c(1L, 3L, 4L), sum, na.rm = TRUE)
    
    apply(tab, c(1L, 2L), computeBookingRate)
  }
}
```

#### Mostly-Harmless Misdemeanors

We begin by examining the booking rates for juveniles arrested for less serious offenses, which we would expect to be very low if the data are accurate and comprehensive. 

```{r 'analysis_bookingRates_misdemeanor_booking_rates_def', echo = FALSE}
if (runAnalysis) {
  macr.juv <- subset(macr.sub, age < 18)
  
  offenseCodes.juv <- c(31, 32, 34, 44, 46)
  offenseCodeNames.juv <- getNameForSummaryOffenseCode(offenseCodes.juv)
  
  br.juv.yearly <-
    getJurisdictionalBookingRatesForOffenseCodeSetByYear(macr.juv, offenseCodes.juv)

  highJurisdictions <-
    apply(br.juv.yearly, 1L, function(row) sum(row > 0.99, na.rm = TRUE) >= 2L)
  
  juvenileOffenseCodes <- paste0(paste0(head(offenseCodeNames.juv, -1), collapse = ", "), ", and ", tail(offenseCodeNames.juv, 1))
  numHighJurisdictions <- sum(highJurisdictions)
   
  writeLines(c(juvenileOffenseCodes, as.character(numHighJurisdictions)),
             file.path("..", txtPath, "analysis_bookingRates_juvenileMisdemeanors_1.txt"))
} else {
  lines <- readLines(file.path("..", txtPath, "analysis_bookingRates_juvenileMisdemeanors_1.txt"))
  juvenileOffenseCodes <- lines[1L]
  numHighJurisdictions <- as.integer(lines[2L])
  rm(lines)
}
```

The codes are: `r juvenileOffenseCodes`. One immediate cause for concern is the number of jursidictions that, for some years, have reported booking rates close to 100%. Each line in the figure below represents a separate law enforcement agency's yearly booking rate beginning in 2005. The figure includes the,`r numHighJurisdictions` jursidictions that have a booking rate in excess of 0.99 for at least two years. 

```{r 'analysis_bookingRates_misdemeanor_booking_rates_1_plot', echo = FALSE}
if (runAnalysis) {
  pdf(file.path("..", imgPath, "analysis_bookingRates_juvenileMisdemeanors_1.pdf"), 6, 7.76)
  plotBookingRateGrid(br.juv.yearly[highJurisdictions,], gridRatio)
  invisible(dev.off())
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_juvenileMisdemeanors_1.pdf"))`

Although a 100% booking rate is implausible, some smaller jurisdictions may have very high booking rates due to small sample sizes. Arbitrarily, we proceed by excluding those that approach 100% three or more times. Another simple cut can be made by excluding jursidictions that are mostly missing, such as those for which have only one recorded value.

```{r 'analysis_bookingRates_misdemeanor_booking_rates_1_exclude', echo = FALSE}
if (runAnalysis) {
  highJurisdictions <-
    apply(br.juv.yearly, 1L, function(row) sum(row > 0.99, na.rm = TRUE) >= 3L)
  br.juv.yearly <- br.juv.yearly[!highJurisdictions,]
  
  missingJurisdictions <-
    apply(br.juv.yearly, 1L, function(row) sum(is.na(row)) >= (numYears - 1L))
  br.juv.yearly <- br.juv.yearly[!missingJurisdictions,]
  
  macr.juv <- subset(macr.juv, ncic_jurisdiction %in% rownames(br.juv.yearly))
  macr.juv$ncic_jurisdiction <- droplevels(macr.juv$ncic_jurisdiction)
}
```

Next, we average each jurisdiction's arrests of juveniles for misdemeanors across years and look at the empirical distribution of booking rates.

```{r 'analysis_bookingRates_misdemeanor_booking_rates_yearly_aggregate', echo = FALSE}
if (runAnalysis) {
  br.juv.agg <- 
    getJurisdictionalBookingRatesForOffenseCodeSet(macr.juv, offenseCodes.juv)
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_juvenileMisdemeanors_2.pdf"), 4.5, 4.5)
  par(mar = c(2.2, 2.2, 2.0, 0.1), mgp = c(1.2, 0.3, 0.0), tcl = -0.5)
  hist(br.juv.agg, main = "Hist. Juv. Misd., 2005+",
       xlab = "Booking Rate", breaks = 20)
  invisible(dev.off())
  
  highJurisdictions <- br.juv.agg >= 0.65
  
  numHighJurisdictions <- sum(highJurisdictions)
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_juvenileMisdemeanors_3.pdf"), 6, 7.76)
  plotBookingRateGrid(br.juv.yearly[highJurisdictions,], gridRatio)
  invisible(dev.off())
  
  writeLines(as.character(numHighJurisdictions),
             file.path("..", txtPath, "analysis_bookingRates_juvenileMisdemeanors_2.txt"))
  
  jurisdictionsToKeep <- names(br.juv.agg)[!highJurisdictions]
  
  rm(highJurisdictions, missingJurisdictions)
  invisible(gc(FALSE))
} else {
  numHighJurisdictions <- as.integer(readLines(file.path("..", txtPath, "analysis_bookingRates_juvenileMisdemeanors_2.txt")))
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_juvenileMisdemeanors_2.pdf"))`

If we assume that the data are relatively complete, there appears to be a plausible cutoff around 0.65, which includes `r numHighJurisdictions` jurisdictions. Their yearly booking rates appear as:

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_juvenileMisdemeanors_3.pdf"))`

While some of these may certainly be valid, we remove them from further analysis because their booking rates are suspiciously high and most appear to have too much year-to-year variation in their booking rates.

#### Serious Felonies

We repeat the above on the reduced set of jurisdictions, but instead of looking at arrest rates for serious felonies. Our goal is to identify jurisdictions that report implausibly low booking rates for serious felonies. 

```{r 'analysis_bookingRates_felony_booking_rates_def', echo = FALSE}
if (runAnalysis) {
  macr.adult <- subset(macr.sub, age >= 18 & ncic_jurisdiction %in% jurisdictionsToKeep)
  macr.adult$ncic_jurisdiction <- droplevels(macr.adult$ncic_jurisdiction)
  
  offenseCodes.adult <- c(6, 14, 25, 8, 9, 12, 27)
  offenseCodeNames.adult <- getNameForSummaryOffenseCode(offenseCodes.adult)
  
  adultOffenseCodes <- paste0(paste0(head(offenseCodeNames.adult, -1), collapse = ", "), ", and ", tail(offenseCodeNames.adult, 1))
  
  br.adult.yearly <-
    getJurisdictionalBookingRatesForOffenseCodeSetByYear(macr.adult, offenseCodes.adult)
  
  lowJurisdictions <-
    apply(br.adult.yearly, 1L, function(row) sum(row < 0.8, na.rm = TRUE) >= 2L)
  
  numLowJurisdictions <- sum(lowJurisdictions)
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_adultFelonies_1.pdf"), 6, 7.76)
  plotBookingRateGrid(br.adult.yearly[lowJurisdictions,], gridRatio)
  invisible(dev.off())
  
  writeLines(c(adultOffenseCodes, as.character(numLowJurisdictions)),
             file.path("..", txtPath, "analysis_bookingRates_adultFelonies.txt"))
} else {
  lines <- readLines(file.path("..", txtPath, "analysis_bookingRates_adultFelonies.txt"))
  adultOffenseCodes <- lines[1L]
  numLowJurisdictions <- as.integer(lines[2L])
  rm(lines)
}
```

The codes correspond to: `r adultOffenseCodes`. Because the rates for these felonies should be high, we highlight any jursidiction that, for two or more years, had a booking rate lower than 80%. This includes `r numLowJurisdictions` jursidictions for which booking rates from 2005 onward are illustrated below:

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_adultFelonies_1.pdf"))`

For comparison, the typical jurisdiction has a straight line near 100%. Some of these look categorically low, including the set in the middle which all decline and increase with each other; others may reflect the variation associated with small sample sizes. For that reason, we set the threshold at three years. Again we look at booking rates across years.

```{r 'analysis_bookingRates_felony_booking_rates_80_exclude', echo = FALSE}
if (runAnalysis) {
  lowJurisdictions <-
    apply(br.adult.yearly, 1L, function(row) sum(row < 0.8, na.rm = TRUE) >= 3L)
  br.adult.yearly <- br.adult.yearly[!lowJurisdictions,]
  
  macr.adult <- subset(macr.adult, ncic_jurisdiction %in% rownames(br.adult.yearly))
  macr.adult$ncic_jurisdiction <- droplevels(macr.adult$ncic_jurisdiction)
  
  jurisdictionsToKeep <- levels(macr.adult$ncic_jurisdiction)
  
  br.adult.agg <- 
    getJurisdictionalBookingRatesForOffenseCodeSet(macr.adult, offenseCodes.adult)
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_adultFelonies_2.pdf"), 4.5, 4.5)
  par(mar = c(2.2, 2.2, 2.0, 0.1), mgp = c(1.2, 0.3, 0.0), tcl = -0.5)
  hist(br.adult.agg, main = "Hist. Adult Felonies 2005+",
       xlab = "Booking Rate", breaks = 20)
  invisible(dev.off())
  
  numRemainingJurisdictions <- nrow(br.adult.yearly)
  writeLines(as.character(numRemainingJurisdictions), 
             file.path("..", txtPath, "analysis_bookingRates_remaining.txt"))
  
  rm(lowJurisdictions)
  invisible(gc(FALSE))
} else {
  numRemainingJurisdictions <- as.integer(readLines(file.path("..", txtPath, "analysis_bookingRates_remaining.txt")))
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_adultFelonies_2.pdf"))`

### Matched ACHS-MACR data

We obtained a probabilistic record linkage for 2014 and 2015 ACHS and MACR data. We use the linked records to estimate the percentage of arrests reported as booked (from MACR) that were actually booked by department (from ACHS). In other words, we use a simple metric: the number of actual bookings (MACR records linked to ACHS) divided by the number of reported bookings (total bookings in MACR). 

We do not possess a full count of individuals who appear in ACHS. This would have allowed us to directly compare the total number of reported bookings to the total number of actual bookings. We would also be able to see what percentage of people who were booked did not appear in the MACR (though DOJ CJIS retroactively fills some of these in). The data we use here contain only the records that could plausibly be linked between MACR and ACHS. The total number of matches was about 374,000 and there were about 2M reported bookings over the same period (out of 2.9M arrests). We believe the matching process misses many true matches (a high false negative rate), which means that we expect that even departments that are accurately reporting bookings would fall well below 100%. 
To examine the quality of booking data, we primarily focus on deparments and offenses. Departmental practices and systems may be responsible for the seemingly small percentage of MACR reported bookings that show up in ACHS. Among the 75% of departments from which we would expect enough bookings to compute statistics (>95 felony arrests), we find large variation in the fraction of reported bookings that could be matched to actual bookings. Apart from deparments reporting bookings incorrectly, this could also be due to department-level variation in the quality of personal data reported on the MACR. If some departments reported poor name and date of birth data, we would find fewer corresponding records in ACHS. 

```{r 'analysis_bookingRates_booking_percent_match_bySumOffenseCode', echo = FALSE}
if (runAnalysis) {
  intersection <- read.csv(file.path("data", "intersection.csv"), stringsAsFactors = FALSE)
  intersection.macr <- intersection[,"MACR_Record", drop = FALSE]
  
  macr.match <- subset(macr, arrest_year >= 2014 & arrest_year <= 2015)
  macr.match$ncic_jurisdiction <- droplevels(macr.match$ncic_jurisdiction)
  macr.match$CID <- seq_len(nrow(macr.match))
  
  denominators <- macr.match
  denominators$matched <- macr.match$CID %in% intersection$MACR_Record
  
  byCrimeTot  <- aggregate(denominators$matched, by = denominators[,"bcs_summary_offense_code", drop = FALSE], length)
  
  denominators$matched[is.na(denominators$matched)] <- FALSE
  byCrimeMatch <- aggregate(denominators$matched == TRUE, by = denominators[,"bcs_summary_offense_code", drop = FALSE], sum)

  bookedCrimes <- merge(byCrimeTot, byCrimeMatch, by = "bcs_summary_offense_code", all.x = TRUE)
  colnames(bookedCrimes) <- c("bcs_summary_offense_code", "total", "matched")
  bookedCrimes$propByCrime <- with(bookedCrimes, matched / total)
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_bookingProportions.pdf"), 4.5, 4.5)
  with(bookedCrimes, {
    n <- total
    x <- matched
    p <- x / n
    
    ## uses Jeffrey's intervals; probably could do something more Bayesian to share
    ## strength across the various categories
    lower <- qbeta(0.025, x + 0.5, n - x + 0.5)
    upper <- qbeta(0.975, x + 0.5, n - x + 0.5)
    
    plot(bcs_summary_offense_code, p,
         ylim = range(lower, p, upper, na.rm = TRUE),
         xlab = "summary offense code",
         ylab = "percentage matched",
         main = "MACR-ACHS bookings match by offense")
    plotOrder <- order(levels(bcs_summary_offense_code))
    lines(rbind(plotOrder, plotOrder, NA),
          rbind(lower, upper, NA))
  })
  invisible(dev.off())
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_bookingProportions.pdf"))`

```{r 'analysis_bookingRates_booking_percent_match_2', echo = FALSE}
if (runAnalysis) {
  jurisd.n <- read.csv(file.path("data", "matchedRecs.csv"))
  jurisd.n <- jurisd.n[,c("ncic_jurisdiction", "summary_offense_level", "bcs_summary_offense_code", "status_type", "disposition")]
  
  ## correct for coding without leading "0"
  levels(jurisd.n$ncic_jurisdiction) <- sapply(levels(jurisd.n$ncic_jurisdiction), function(levelName) if (nchar(levelName) == 3L) paste0("0", levelName) else levelName)

  jurisd <- cbind(ftable(summary_offense_level ~ ncic_jurisdiction, macr.match),
                  ftable(status_type           ~ ncic_jurisdiction, macr.match),
                  ftable(disposition           ~ ncic_jurisdiction, macr.match))
  colnames(jurisd) <- with(macr.match, c(levels(summary_offense_level), levels(status_type), levels(disposition)))
  jurisd <- as.data.frame(jurisd)
  jurisd$ncic_jurisdiction <- levels(macr.match$ncic_jurisdiction)
  jurisd <- jurisd[,c("ncic_jurisdiction", setdiff(colnames(jurisd), "ncic_jurisdiction"))]
  
  jurisd <- jurisd[,c("ncic_jurisdiction", "felony", "misdemeanor", "cited", "booked", "felony complaint sought", "misdemeanor complaint sought")]
  colnames(jurisd) <- c("ncic_jurisdiction", "felArrestTot", "misdArrestTot", "repCiteTot", "repBookTot", "repFelRefTot", "repMisdRefTot")
  
  jurisd$repBothRefTot <- jurisd$repFelRefTot + jurisd$repMisdRefTot
  
  actualBookTot <- aggregate(jurisd.n$status_type, by = list(jurisd.n$ncic_jurisdiction), length)
  jurisd$actualBookTot <- actualBookTot[match(jurisd$ncic_jurisdiction, actualBookTot[,1L]), "x"]
  
  #quantile(jurisd$felArrestTotal)
  #75% of depts. have 95 or more felony arrests. We use these to analyze matched booking rates.  
  stat.jurisd <- subset(jurisd, felArrestTot >= 95)
}
``` 
One item that gives us some condfidence that the booking data contain useful signal is their scatterplots versus other variables on the MACR. We observe the relationships that we would expect. 

```{r 'booking_percent_match_3', echo = FALSE}
if (runAnalysis) {
  pdf(file.path("..", imgPath, "analysis_bookingRates_logStat.pdf"), 6, 6)
  plot(log(stat.jurisd[,2:9]), main = "scatterplots of MACR variables")
  invisible(dev.off())

  stat.jurisd$bookMatchPct <- stat.jurisd$actualBookTot / stat.jurisd$repBookTot
  # there are two departments with matched booking rates higher than 1 (3623 and 3702). One had 271 fel arrests, the other 3031. This suggeststhat rather than over-reporting bookings, they are under-reporting bookings. Along with some departments that have 100% or nearly 100% match rates, we suspect these departments had their MACR data created from the ACHS data by CJIS. 
#stat.jurisd$ncic_jurisdiction[stat.jurisd$bookMatchPct > 1]
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_bookedMatchedDensity.pdf"), 4.5, 4.5)
  plot(density(stat.jurisd$bookMatchPct), xlim=c(0,1), main="distribution of department booking match averages")
  invisible(dev.off())
  # the 5th percentile is 9.1% and the 95th percentile on matched records is 31.0%
  #quantile(stat.jurisd$bookMatchPct, .05)
  #quantile(stat.jurisd$bookMatchPct, .95)

  stat.jurisd$actualBookvsRepRefs <- stat.jurisd$repBothRefTot / stat.jurisd$actualBookTot 
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_actualBookedHist.pdf"), 4.5, 4.5)
  hist(stat.jurisd$actualBookvsRepRefs, main="department ratios of reported to matched bookings", breaks=50, xlab="ratio")
  invisible(dev.off())

  #an alternative metric is to compare the number of reported felony arrests with the number of matched bookings
  stat.jurisd$repFelArrest_actualBook <- stat.jurisd$felArrestTot / stat.jurisd$repBookTot 
  #quantile(stat.jurisd$repFelArrest_actualBook, .05)
  #quantile(stat.jurisd$repFelArrest_actualBook, .95)
  
  #we see very little relationship between total arrests (dept. size) and booking matches
  pdf(file.path("..", imgPath, "analysis_bookingRates_repBookedVDiff.pdf"), 4.5, 4.5)
  plot(log(stat.jurisd$repBookTot), stat.jurisd$actualBookTot/stat.jurisd$repBookTot,
       ylim = c(0, 1),
       main = "proportion of matched bookings by arrest volume",
       xlab = "ln(department arrests)",
       ylab = "proportion of reported bookings matched")
  lines(loess.smooth(log(stat.jurisd$repBookTot), stat.jurisd$actualBookTot/stat.jurisd$repBookTot, ylim=c(0,1)))
  invisible(dev.off())
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_logStat.pdf"))`

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_bookedMatchedDensity.pdf"))`

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_actualBookedHist.pdf"))`

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_repBookedVDiff.pdf"))`

### Final Visualization

At this point, the `r numRemainingJurisdictions` jurisdictions that remain seem to be adequate in terms of variabilty. Their booking rates for juvenile misdemeanors are:

```{r 'analysis_bookingRates_misdemeanor_booking_rates_plot', echo = FALSE}
if (runAnalysis) {
  br.juv.yearly <- br.juv.yearly[rownames(br.juv.yearly) %in% rownames(br.adult.yearly),]
  
  pdf(file.path("..", imgPath, "analysis_bookingRates_remaining_1.pdf"), 6, 7.76)
  plotBookingRateGrid(br.juv.yearly, gridRatio)
  invisible(dev.off())
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_remaining_1.pdf"))`

The yearly booking rates for serious felonies are:

```{r 'analysis_bookingRates_felony_booking_rates_plot', echo = FALSE}
if (runAnalysis) {
  pdf(file.path("..", imgPath, "analysis_bookingRates_remaining_2.pdf"), 6, 7.76)
  plotBookingRateGrid(br.adult.yearly, gridRatio)
  invisible(dev.off())
}
```

`r rmdImageInline(file.path(imgPath, "analysis_bookingRates_remaining_2.pdf"))`

## Booking Rate Tables

```{r 'analysis_bookingRates_booking_rate_function', echo = FALSE}
if (runAnalysis) getOffenseCodeBookingRatesForJurisdictions <- function(macr, jurisdictions)
{
  macr.sub <- macr[,c("bcs_summary_offense_code", "status_type",
                      "race_or_ethnicity", "offense_level")]
  macr.sub <- subset(macr.sub, macr$ncic_jurisdiction %in% jurisdictions)
  offense_level <- macr.sub$offense_level
  macr.sub$offense_level <- NULL
  
  macr.fel <- subset(macr.sub, offense_level == "felony")
  macr.mis <- subset(macr.sub, offense_level == "misdemeanor")
  
  tableBookingRates <- function(macr) {
    getBookingRatesFromTable <- function(tab) {
      ## change the result here to change what ends up in the big
      ## table at the end
      N <- rowSums(tab, na.rm = TRUE)
      data.frame(N = N,
                 bkd = tab[,"booked"] / N,
                 ctd  = tab[,"cited"] / N)
      #data.frame(bkd = tab[,"booked"],
      #           ctd  = tab[,"cited"],
      #           "bkd%" = tab[,"booked"] / rowSums(tab, na.rm = TRUE))
    }
    tab <- table(macr)
    
    totals <- apply(tab, c(1, 2), sum)
    
    br.all   <- getBookingRatesFromTable(apply(tab, c(1, 2), sum))
    br.white <- getBookingRatesFromTable(tab[,,"White"])
    br.black <- getBookingRatesFromTable(tab[,,"Black"])
    br.hisp  <- getBookingRatesFromTable(tab[,,"Hispanic"])
    
    result <- cbind(br.all, br.white, br.black, br.hisp)
    colnames(result) <- c(
      colnames(br.all), 
      paste0(colnames(br.white), ".wt"),
      paste0(colnames(br.black), ".bl"),
      paste0(colnames(br.hisp), ".hs"))
    
    offenseCodes <- as.integer(rownames(result))
    rownames(result) <- paste0(offenseCodes, "-", getNameForSummaryOffenseCode(offenseCodes))
    result
  }
  
  list(felony      = tableBookingRates(macr.fel),
       misdemeanor = tableBookingRates(macr.mis))
}
```

```{r 'analysis_bookingRates_booking_rates', echo = FALSE}
if (runAnalysis) {
  br <- getOffenseCodeBookingRatesForJurisdictions(macr.sub, jurisdictionsToKeep)
  
  ## always sorts using the first column
  br.felony <- br$felony[order(br$felony[,1], decreasing = TRUE),]
  ## cuts off the offense code names to a prefix
  rownames(br.felony) <- substr(rownames(br.felony), 1, 12)
  br.felony <- br.felony[br.felony$N > 0L,]
  
  write.csv(br.felony, file.path("..", txtPath, "analysis_bookingRates_felonyTable.csv"))
  
  br.misdemeanor <- br$misdemeanor[order(br$misdemeanor[,1], decreasing = TRUE),]
  rownames(br.misdemeanor) <- substr(rownames(br.misdemeanor), 1, 12)
  br.misdemeanor <- br.misdemeanor[br.misdemeanor$N > 0L,]
  
  write.csv(br.felony, file.path("..", txtPath, "analysis_bookingRates_misdemeanorTable.csv"))
} else {
  br.felony <- read.csv(file.path("..", txtPath, "analysis_bookingRates_felonyTable.csv"), row.names = 1L)
  br.misdemeanor <- read.csv(file.path("..", txtPath, "analysis_bookingRates_misdemeanorTable.csv"), row.names = 1L)
}
```

`r if (rmdGetOutputFormat() == "latex") "\\blandscape" else ""`

### Juvenile Misdemeanors

```{r echo = FALSE, results = "asis"}
tab <- round(br.misdemeanor, 2)
tab[["offense code"]] <- rownames(tab)
rownames(tab) <- NULL
tab <- tab[,c("offense code", setdiff(colnames(tab), "offense code"))]
cat(nativeFormat(tab, size = "footnotesize", width = 9,
                 colWidths = sapply(seq_along(tab), function(i) max(nchar(tab[[i]]), nchar(colnames(tab)[i])))),
    sep = "\n")
```

`r if (rmdGetOutputFormat() == "latex") "\\elandscape" else ""`

`r if (rmdGetOutputFormat() == "latex") "\\blandscape" else ""`

### Adult Felonies

```{r echo = FALSE, results = "asis"}
tab <- round(br.felony, 2)
tab[["offense code"]] <- rownames(tab)
rownames(tab) <- NULL
tab <- tab[,c("offense code", setdiff(colnames(tab), "offense code"))]
cat(nativeFormat(tab, size = "footnotesize", width = 9,
                 colWidths = sapply(seq_along(tab), function(i) max(nchar(tab[[i]]), nchar(colnames(tab)[i])))),
    sep = "\n")

rm(tab)
```

`r if (rmdGetOutputFormat() == "latex") "\\elandscape" else ""`

To summarize, we know that:

* Due to the high false negative rate in linking records, the match percentage of a department that accurately reports bookings will be below 100%
* Reported booking rates, even for less serious misdemeanors, appear to be too high 
* The percentage of individuals arrested who are on parole or probation (22%) is too low to explain these high booking rates
* The scatterplots comparing bookings to other variables seem reasonable, suggesting there is useful information in the status_type field
* The only hypothesis we have that would explain all these issues is that the MACR reflects under-reporting of arrests that are not booked (typically, those that are less serious), thus shrinking the denominator on booking rates and making it seem that booking rates are very high across all crimes. There may be other explanations and we think it might be worth our calling several different police departments to speak with their records staff. 

```{r echo = FALSE}
invisible(gc(FALSE))
```
