```{r 'clean_setup', echo = FALSE}
cleanDataFile <- file.path("..", dataPath, tables$file[match(mainTable, tables$name)])

txtFiles <- file.path("..", txtPath, c("clean_type_32.txt",
                                       "clean_arrestDate.txt"))

performClean <- !file.exists(cleanDataFile) || any(!file.exists(txtFiles))

if (performClean)
  macr.clean <- loadData(paste0(mainTable, "_dirty"), path = "..")

rm(txtFiles)
```

`r rmdPageBreak()`

# Data Cleaning

## Deleted Records

Records with a type id of 32 represent deleted rows, and as they contain no information about the kind of arrest made are deemed unusuable and omitted. A typical example would be:

```{r 'clean_type_32', echo = FALSE, results = "asis"}
txtFile <- file.path("..", txtPath, "clean_type_32.txt")
if (performClean) {
  badRows <- macr.clean$record_type_id == "32"
  
  type32 <- rmdDisplayDataFrameHead(macr.clean[which(badRows)[seq_len(4L)],], maxCols = 10L, maxColWidth = 5L)
  writeLines(type32, txtFile)
  
  macr.clean <- macr.clean[!badRows,]
  macr.clean$record_type_id <- droplevels(macr.clean$record_type_id)
  
  ## removing rows ends up dropping levels from a few factors
  macr.clean$bcs_jurisdiction  <- droplevels(macr.clean$bcs_jurisdiction)
  macr.clean$ncic_jurisdiction <- droplevels(macr.clean$ncic_jurisdiction)
  
  rm(badRows)
} else {
  type32 <- readLines(txtFile)
}
cat(type32, sep = "\n")
rm(type32, txtFile)
```

## Arrest Date

```{r 'clean_arrestdate', echo = FALSE}
txtFile <- file.path("..", txtPath, "clean_arrestDate.txt")
if (performClean) {
  macr.clean$arrest_date <- with(macr.clean, as.Date(paste(arrest_year, arrest_month, arrest_day, sep = "-"), format = "%F"))
  macr.clean$arrest_day[macr.clean$arrest_day <= 0L] <- NA
  
  naRows <- is.na(macr.clean$arrest_date) & !is.na(macr.clean$arrest_day)
  naIndices <- which(naRows)
  
  ## month rolled over into the next accidentially
  monthSlippedRows <- with(macr.clean, sapply(naIndices, function(index) 
    index > 1L                                     &&
    arrest_month[index - 1L] < arrest_month[index] &&
    arrest_day[index - 1L]  <= arrest_day[index]   &&
    ncic_jurisdiction[index - 1L] == ncic_jurisdiction[index]
  ))
  
  monthSlipped <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[which.max(monthSlippedRows)],
                             c("ncic_jurisdiction", "arrest_year", "arrest_month", "arrest_day", "arrest_date")])
  
  ## if it looks like maybe a day has transposed digits, it'll be the same month as everything
  ## else but higher than its neighbors
  dayTransposedRows <- with(macr.clean, sapply(naIndices, function(index)
    (index == 1L || (arrest_month[index] == arrest_month[index - 1L] &&
                     arrest_day[index]   >  arrest_day[index - 1L]   &&
                     ncic_jurisdiction[index] == ncic_jurisdiction[index - 1L]))        &&
    (index == length(arrest_day) || (arrest_month[index] == arrest_month[index + 1L] &&
                                     arrest_day[index] > arrest_day[index + 1L]     &&
                                     ncic_jurisdiction[index] == ncic_jurisdiction[index + 1L]))
  ))
  
  dayTransposedRow <- { for (i in which(dayTransposedRows))
    if (macr.clean$arrest_day[naIndices[i]] == 31L && all(macr.clean$arrest_day[naIndices[i] + c(-1L, 1L)] == 13L)) break ; i }
  dayTransposed <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[dayTransposedRow],
                             c("ncic_jurisdiction", "arrest_year", "arrest_month", "arrest_day", "arrest_date")])
  
  ## looks like someone just extended february
  dayAddedRows <- with(macr.clean, sapply(naIndices, function(index)
    index > 1L && index < length(arrest_day)               &&
    length(unique(arrest_month[index + -1L:1L])) == 1L     &&
    length(unique(ncic_jurisdiction[index + 1L:1L])) == 1L &&
    all(arrest_day[index + 0L:1L] == (arrest_day[index - 1L] + 1L), na.rm = TRUE)
  ))
  
  dayAdded <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[which.max(dayAddedRows)],
                             c("ncic_jurisdiction", "arrest_year", "arrest_month", "arrest_day", "arrest_date")])
  
  numBadArrestDate <- as.character(sum(naRows))
  arrestDate <- c(numBadArrestDate, monthSlipped, dayTransposed, dayAdded)
  
  writeLines(arrestDate, txtFile)
  
  rm(naRows, naIndices, monthSlippedRows, monthSlipped, dayTransposedRows, i,
     dayTransposed, dayAddedRows, dayAdded, numBadArrestDate)
} else {
  arrestDate <- readLines(txtFile)
}
rm(txtFile)
```

A total of `r arrestDate[1L]` records cannot be parsed into valid dates, i.e. the combination of `arrest_year`, `arrest_month`, and `arrest_day` results in a non-sensical date. Records with `arrest_day` of 0 are changed to `NA`, however in other cases it may the arrest day or month that causes the failure to parse. Some structure in the coding may allow the correct date to be recovered. For example:

* Records at the end of a month with the `arrest_month` field prematurely incremented:

```{r, echo = FALSE, results = "asis"}
cat(arrestDate[seq.int(2L, 18L)], "\n\n", sep = "\n")
```

* Records where the digits in `arrest_day` appear to be transposed:

```{r echo = FALSE, results = "asis"}
cat(arrestDate[seq.int(19L, 35L)], "\n\n", sep = "\n")
```

* Records where days were added at the end of a month:

```{r echo = FALSE, results = "asis"}
cat(arrestDate[seq.int(36L, 52L)], "\n\n", sep = "\n")
rm(arrestDate)
```

As for now determining a 'correct arrest date' requires estimation, we leave those as is and augment the data with a field `arrest_date` that contains `NA` for all rows where a date cannot be parsed.

```{r 'clean_results', echo = FALSE}
if (performClean) {
  macr <- macr.clean
  save(macr, file = cleanDataFile)
  
  rm(macr.clean)
  invisible(gc(FALSE))
}
rm(performClean, cleanDataFile)
```