# Data Cleaning

```{r clean, echo=FALSE}
cleanComplete <- file.exists(file.path(dataPath, "macr_clean_pii.Rdata"))
```

## Deleted Records

Records with a type id of 32 represent deleted rows, and as they contain no information about the kind of arrest made are deemed unusuable and omitted. A typical example would be:

```{r clean_type_32, echo=FALSE, results = "asis"}
cleanFile <- file.path("..", txtPath, "clean_type_32.txt")
if (!file.exists(cleanFile)) {
  macr <- loadData("macr")
  badRows <- macr$record_type_id == 32
  
  type32 <- rmdDisplayDataFrameHead(macr[which(badRows)[seq_len(4L)],], maxCols = 10L, maxColWidth = 5L)
  writeLines(type32, cleanFile)
  
  macr.clean <- macr[!badRows,]
  macr.clean$record_type_id <- droplevels(macr.clean$record_type_id)
  
  rm(badRows)
} else {
  type32 <- readLines(cleanFile)
}
cat(type32, sep = "\n")
rm(type32, cleanFile)
```

## Arrest Date

```{r clean_arrestdate, echo=FALSE}
cleanFile <- file.path("..", txtPath, "clean_arrestDate.txt")
if (!file.exists(cleanFile)) {
  macr.clean$arrest_date <- with(macr.clean, as.Date(paste(arrest_year, arrest_month, arrest_day, sep = "-"), format = "%F"))
  macr.clean$arrest_day[macr.clean$arrest_day == 0L] <- NA
  
  naRows <- is.na(macr.clean$arrest_date) & !is.na(macr.clean$arrest_day)
  naIndices <- which(naRows)
  
  ## month rolled over into the next accidentially
  monthSlippedRows <- with(macr.clean, sapply(naIndices, function(index) 
    index > 1L                                     &&
    arrest_month[index - 1L] < arrest_month[index] &&
    arrest_day[index - 1L]  <= arrest_day[index]   &&
    ncic_jurisdiction[index - 1L] == ncic_jurisdiction[index]
  ))
  
  monthSlipped <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[which.max(monthSlippedRows)],
                             c("arrest_year", "arrest_month", "arrest_day", "arrest_date", "ncic_jurisdiction")])
  
  ## if it looks like maybe a day has transposed digits, it'll be the same month as everything
  ## else but higher than its neighbors
  dayTransposedRows <- with(macr.clean, sapply(naIndices, function(index)
    (index == 1L || (arrest_month[index] == arrest_month[index - 1L] &&
                     arrest_day[index]   >  arrest_day[index - 1L]   &&
                     ncic_jurisdiction[index] == ncic_jurisdiction[index - 1L]))        &&
    (index == length(arrest_day) || (arrest_month[index] == arrest_month[index + 1L] &&
                                     arrest_day[index] > arrest_day[index + 1L]     &&
                                     ncic_jurisdiction[index] == ncic_jurisdiction[index + 1L]))
  ))
  
  dayTransposedRow <- { for (i in which(dayTransposedRows))
    if (macr.clean$arrest_day[naIndices[i]] == 31L && all(macr.clean$arrest_day[naIndices[i] + c(-1L, 1L)] == 13L)) break ; i }
  dayTransposed <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[dayTransposedRow],
                             c("arrest_year", "arrest_month", "arrest_day", "arrest_date", "ncic_jurisdiction")])
  
  ## looks like someone just extended february
  dayAddedRows <- with(macr.clean, sapply(naIndices, function(index)
    index > 1L && index < length(arrest_day)               &&
    length(unique(arrest_month[index + -1L:1L])) == 1L     &&
    length(unique(ncic_jurisdiction[index + 1L:1L])) == 1L &&
    all(arrest_day[index + 0L:1L] == (arrest_day[index - 1L] + 1L), na.rm = TRUE)
  ))
  
  dayAdded <- rmdFormat(macr.clean[0L:6L - 3L + naIndices[which.max(dayAddedRows)],
                             c("arrest_year", "arrest_month", "arrest_day", "arrest_date", "ncic_jurisdiction")])
  
  numBadArrestDate <- as.character(sum(naRows))
  arrestDate <- c(numBadArrestDate, monthSlipped, dayTransposed, dayAdded)
  
  writeLines(arrestDate, cleanFile)
  
  #macr.clean$arrest_month[naRows] <- NA
  #macr.clean$arrest_day[naRows]   <- NA
  
  rm(naRows, naIndices, monthSlippedRows, monthSlipped, dayTransposedRows, i,
     dayTransposed, dayAddedRows, dayAdded, numBadArrestDate)
} else {
  arrestDate <- readLines(cleanFile)
}
rm(cleanFile)
```

A total of `r arrestDate[1L]` records cannot be parsed into valid dates, i.e. the combination of `arrest_year`, `arrest_month`, and `arrest_day` results in a non-sensical date. Record with `arrest_day` of 0 are straightforward to handle, while in other cases it may be either the day or month that has been recorded incorrectly. Some structure in the coding may allow the correct date to be recovered. For example:

* Records at the end of a month with the `arrest_month` field prematurely incremented:

```{r, echo=FALSE, results="asis"}
cat(arrestDate[seq.int(2L, 10L)], "\n\n", sep = "\n")
```
* Records where the digits in `arrest_day` appear to be transposed:

```{r echo=FALSE, results="asis"}
cat(arrestDate[seq.int(11L, 19L)], "\n\n", sep = "\n")
```
* Records where days were added at the end of a month:

```{r echo=FALSE, results="asis"}
cat(arrestDate[seq.int(20L, 28L)], "\n\n", sep = "\n")
rm(arrestDate)
```

As for now determining a 'correct arrest date' requires estimation, we leave those as is and augment the data with a field `arrest_date` that contains `NA` for all rows where a date cannot be parsed.

##  Birth Date

MACR includes fields for both the birth date and age, which do not always align. Errors in the birth date include:

* birth year recorded as `19xx` instead of `18xx`
* birth year recorded as `9xx` instead of `19xx`
* birth year in wrong millenia
* birth day 0
* birth month 0
* birth date invalid (e.g. `02/31/1991`)

```{r clean_birthdate_1, echo=FALSE}
cleanFile <- file.path("..", txtPath, "clean_birthdate_1.txt")
if (!file.exists(cleanFile)) {
  birth_year.clean <- macr.clean$birth_year

  discrepancyRows <- with(macr.clean, abs(arrest_year - birth_year - age) > 1L)
  discrepancyRows[is.na(discrepancyRows)] <- FALSE
  numDiscrepancyBirthYear <- as.character(sum(discrepancyRows))
  
  offBy100 <- discrepancyRows & with(macr.clean, abs(arrest_year - (birth_year - 100L) - age) <= 1L)
  offBy100[is.na(offBy100)] <- FALSE
  
  birth_year.clean[offBy100] <- birth_year.clean[offBy100] - 100L
  
  discrepancyRows <- discrepancyRows & !offBy100
  
  offBy1000 <- discrepancyRows & with(macr.clean, abs(arrest_year - (birth_year + 1000L) - age) <= 1L)
  offBy1000[is.na(offBy1000)] <- FALSE

  birth_year.clean[offBy1000] <- birth_year.clean[offBy1000] + 1000L
  
  discrepancyRows <- discrepancyRows & !offBy1000
  
  birth_year.clean[which(birth_year.clean == 3082L)] <- 1982L
  birth_year.clean[which(birth_year.clean == 5010L)] <- 1910L
  
  macr.clean$birth_year <- birth_year.clean
  rm(discrepancyRows, offBy100, offBy1000, birth_year.clean)
  
  writeLines(numDiscrepancyBirthYear, cleanFile)
} else {
  numDiscrepancyBirthYear <- readLines(cleanFile)
}
rm(cleanFile)
```

Errors in the birth year are detected by looking for when the age and distance from arrest year to birth year exceed 1 in absolute value, and are `r numDiscrepancyBirthYear` in number. Cases where there difference is 100 or 1000 years can be resolved automatically, which leaves only 2 cases to be resolved by hand.

```{r clean_birthdate_2, echo=FALSE}
cleanFile <- file.path("..", txtPath, "clean_birthdate_2.txt")
if (!file.exists(cleanFile)) {
  macr.clean$birth_month[macr.clean$birth_month <= 0L] <- NA
  macr.clean$birth_day  [macr.clean$birth_day   <= 0L] <- NA
  
  macr.clean$birth_date <- with(macr.clean, as.Date(paste(birth_year,  birth_month,  birth_day,  sep = "-"), format = "%F"))
  
  ## load getTimeDifferenceInYears
  source(file.path("..", commonSrcPath, "time.R"))
  age.est <- with(macr.clean, getTimeDifferenceInYears(arrest_date, birth_date))
  discrepancyRows <- macr.clean$age != age.est
  discrepancyRows[is.na(discrepancyRows)] <- FALSE
  
  numDiscrepancyBirthYear <- sum(discrepancyRows)
  rm(age.est, discrepancyRows, getTimeDifferenceInYears)
  
  writeLines(as.character(numDiscrepancyBirthYear), cleanFile)
} else {
  numDiscrepancyBirthYear <- as.integer(readLines(cleanFile))
}
rm(cleanFile)
```

Of the remaining cases where a valid birth date can be derived from the birth year, month, and day, there `r pl <- numDiscrepancyBirthYear != 1L; paste(if (pl) "are" else "is", numDiscrepancyBirthYear, if (pl == 1L) "cases" else "case")` where the distance from arrest date to birth date does not yield the recorded age. `r colFmt("red", "TODO: transition")`

```{r clean_birthdate_3, echo=FALSE}
rm(numDiscrepancyBirthYear)
cleanTxtFile <- file.path("..", txtPath, "clean_birthdate_3.txt")
cleanImgFile <- file.path("..", imgPath, "clean_birthmonthHistogram.pdf")
if (!file.exists(cleanTxtFile) || !file.exists(cleanImgFile)) {
  naRows <- with(macr.clean, is.na(birth_date) & !is.na(birth_month) & !is.na(birth_day))
  
  numInvalidBirthDates       <- sum(naRows)
  numInvalidBirthDates.feb30 <- with(macr.clean, sum(birth_day[naRows] == 30L & birth_month[naRows] == 2L))
  writeLines(as.character(c(numInvalidBirthDates, numInvalidBirthDates.feb30)), cleanFile)
  
  pdf(cleanImgFile, 3, 3)
  with(macr.clean,
       discrete.histogram(birth_month[naRows & (birth_day != 30L | birth_month != 2L)],
                          xlab = "Month", col = "gray", border = NA, main = "Birth Month, Not 02/30"))
  dev.off()
  
} else {
  lines <- readLines(cleanTxtFile)
  numInvalidBirthDates <- as.integer(lines[1L])
  numInvalidBirthDates.feb30 <- as.integer(lines[2L])
  rm(lines)
}
rm(cleanTxtFile, cleanImgFile)
```

The MACR manual states that if the specific birth date is unknown, the birth month and day should be recorded as February 30th. This accounts for `r numInvalidBirthDates.feb30` of the remaining `r numInvalidBirthDates` records with an invalid birth date. After excluding these records, there are still an anomalous amount of birth dates in February.

```{r clean_birthdate_4, echo=FALSE}
imgText <- paste0("![](", file.path(imgPath, "clean_birthmonthHistogram.pdf"), ")")
cat(imgText, "\n\n", sep = "")
rm(imgText, numInvalidBirthDates, numInvalidBirthDates.feb30)
```

## Age

```{r clean_ageHistogram, echo=FALSE, results="asis"}
ageHistogramFile <- file.path("..", imgPath, "clean_ageHistogram.pdf")
if (!file.exists(ageHistogramFile)) {
  tab <- table(macr.clean$age)
  ages <- as.integer(names(tab))
  pdf(ageHistogramFile, width = 6, height = 3)
  par(mfrow = c(1L, 2L))
  discrete.histogram(tab[ages <= 10], xlab = "age", col = "gray", main = "age <= 10")
  discrete.histogram(tab[ages >= 80], xlab = "age", col = "gray", main = "age >= 80")
  dev.off()
  rm(tab, ages)
}
ageHistogramFile <- file.path(imgPath, "clean_ageHistogram.pdf")
imgText <- paste0("![](", ageHistogramFile, ")")

cat(imgText, "\n\n", sep = "")
rm(imgText, ageHistogramFile)
```

As the above figure shows, there are a large number of arrests for exceptionally young children, and an odd decrease in arrests at age 90. For the young we find:

```{r clean_youngTable, echo=FALSE, results="asis"}
cleanFile <- file.path("..", txtPath, "clean_youngTable.txt")
if (!file.exists(cleanFile)) {
  youngTable <- rmdFormat(table(macr.clean[macr.clean$age <= 5, c("offense_level", "age")]))
  writeLines(youngTable, cleanFile)
} else {
  youngTable <- readLines(cleanFile)
}

cat(youngTable, sep = "\n")
rm(youngTable, cleanFile)
```